---
title: "NetworkProperties"
author: "Ina Maria Deutschmann"
date: "9/11/2022"
output: html_document
---

# Network properties
set path
```{r}
PATH <- "~/Documents/Ina/Pedro/"
```

Prerequisites
```{r, echo = FALSE}
library(igraph)
library(ggplot2)
library(ggridges)
theme_set(theme_minimal())
library(reshape2)
Color_Months <- c(Jan="#08519C", Feb="#2171B5", Mar="#6BAED6",
                  Apr="#006D2C", May="#41AB5D", Jun="#74C476",
                  Jul="#980043", Aug="#CE1256", Sep="#DF65B0",
                  Oct="#993404", Nov="#CC4C02", Dec="#EC7014")
```

function: get_nw_properties(dt)
```{r}
get_nw_proterties <- function(dt, col_associationScore, ID){
  parameters <- c('num_nodes', 'num_edges', 'edge_density', 'num_pos_edges', 'portion_pos_edges', 'num_neg_edges', 'portion_neg_edges',
                  'average_pos_association_strength','average_neg_association_strength',
                  'transitivity_global_clusering_coef', 'transitivity_ave_clustering_coef_mean_local',
                  'ave_path_length',
                  'assortativity_nominal_EukProk', 'assortativity_nominal_Sizefraction', 'assortativity_degree',
                  'count_nano', 'count_pico', 'count_euk', 'count_prok', 'count_euk_nano', 'count_euk_pico', 'count_prok_nano', 'count_prok_pico', 
                  'fraction_nano', 'fraction_pico', 'fraction_euk', 'fraction_prok', 'fraction_euk_nano', 'fraction_euk_pico', 'fraction_prok_nano', 'fraction_prok_pico')
  NWpropOverview <- data.frame(newNW = rep(0, length(parameters)))
  rownames(NWpropOverview) <- parameters
  # new entry
  NWpropOverview$newNW
  # Create igraph object
  G <- graph_from_data_frame(dt, directed = FALSE)
  # number of nodes without isolated nodes
  NWpropOverview['num_nodes', 'newNW'] <- vcount(G)
  # number of edges (`num_edges`) 
  NWpropOverview['num_edges', 'newNW'] <- ecount(G)
  # edge density (`edge_density`): The proportion of present edges from all possible edges in the network.
  NWpropOverview['edge_density', 'newNW'] <- edge_density(simplify(G), loops = FALSE)
  # number of positive edges (`num_pos_edges`)
  NWpropOverview['num_pos_edges', 'newNW'] <- sum(dt[,col_associationScore]>0)
  # percentage of positive edges (`portion_pos_edges`)
  NWpropOverview['portion_pos_edges', 'newNW'] <- NWpropOverview['num_pos_edges', 'newNW'] / NWpropOverview['num_edges', 'newNW']
  # number of negative edges (`num_neg_edges`)
  NWpropOverview['num_neg_edges', 'newNW'] <- sum(dt[,col_associationScore]<0)
  # percentage of negative edges (`portion_neg_edges`)
  NWpropOverview['portion_neg_edges', 'newNW'] <- NWpropOverview['num_neg_edges', 'newNW'] / NWpropOverview['num_edges', 'newNW']
  # average positive association strength (`average_pos_association_strength`)
  NWpropOverview['average_pos_association_strength', 'newNW'] <- mean(dt[dt[,col_associationScore]>0,col_associationScore])
  # average negative association strength (`average_neg_association_strength`)
  NWpropOverview['average_neg_association_strength', 'newNW'] <- mean(dt[dt[,col_associationScore]<0,col_associationScore])
  # transitivity of a graph
  NWpropOverview['transitivity_global_clusering_coef', 'newNW'] <- transitivity(G, type="global")
  NWpropOverview['transitivity_ave_clustering_coef_mean_local', 'newNW'] <- transitivity(G, type="average")
  # average path length (`ave_path_length`): Average path length: the mean of the shortest distance between each pair of nodes in the network (in both directions for directed graphs).
  NWpropOverview['ave_path_length', 'newNW'] <- mean_distance(G, directed = FALSE, unconnected = TRUE)
  # assortativity_nominal_EukProk
  types <- rep(NA, length(V(G)$name))  # Initialize a vector with NA
  types[grep("_18S_", V(G)$name)] <- 1  # Assign 1 for nodes containing 18S
  types[grep("_16S_", V(G)$name)] <- 2  # Assign 2 for nodes containing 16S
  NWpropOverview['assortativity_nominal_EukProk', 'newNW'] <- assortativity_nominal(G, types, directed = FALSE)
  # assortativity_nominal_Sizefraction
  types <- rep(NA, length(V(G)$name))  # Initialize a vector with NA
  types[grep("S_022_3", V(G)$name)] <- 1  # Assign 1 for nodes containing small size fraction
  types[grep("S_3_200", V(G)$name)] <- 2  # Assign 2 for nodes containing big size fraction
  NWpropOverview['assortativity_nominal_Sizefraction', 'newNW'] <- assortativity_nominal(G, types, directed = FALSE)
  # assortativity_degree
  NWpropOverview['assortativity_degree', 'newNW'] <- assortativity_degree(G, directed = FALSE)
  
  # count and fraction of Euk/Prok and nano/pico
  Nodes <- V(G)$name
  NWpropOverview['count_nano', 'newNW'] <- length(grep("3_200",Nodes))
  NWpropOverview['count_pico', 'newNW'] <- length(grep("022_3",Nodes))
  
  NWpropOverview['count_euk', 'newNW'] <- length(grep("18S",Nodes))
  NWpropOverview['count_prok', 'newNW'] <- length(grep("16S",Nodes))
  
  NWpropOverview['count_euk_nano', 'newNW'] <- length(grep("18S_3_200",Nodes))
  NWpropOverview['count_euk_pico', 'newNW'] <- length(grep("18S_022_3",Nodes))
  
  NWpropOverview['count_prok_nano', 'newNW'] <- length(grep("16S_3_200",Nodes))
  NWpropOverview['count_prok_pico', 'newNW'] <- length(grep("16S_022_3",Nodes))
  
  NWpropOverview['fraction_nano', 'newNW'] <- length(grep("3_200",Nodes))/length(Nodes)*100
  NWpropOverview['fraction_pico', 'newNW'] <- length(grep("022_3",Nodes))/length(Nodes)*100
  
  NWpropOverview['fraction_euk', 'newNW'] <- length(grep("18S",Nodes))/length(Nodes)*100
  NWpropOverview['fraction_prok', 'newNW'] <- length(grep("16S",Nodes))/length(Nodes)*100
  
  NWpropOverview['fraction_euk_nano', 'newNW'] <- length(grep("18S_3_200",Nodes))/length(Nodes)*100
  NWpropOverview['fraction_euk_pico', 'newNW'] <- length(grep("18S_022_3",Nodes))/length(Nodes)*100
  
  NWpropOverview['fraction_prok_nano', 'newNW'] <- length(grep("16S_3_200",Nodes))/length(Nodes)*100
  NWpropOverview['fraction_prok_pico', 'newNW'] <- length(grep("16S_022_3",Nodes))/length(Nodes)*100
  
  # update ID
  colnames(NWpropOverview)[which(colnames(NWpropOverview)=="newNW")] <- ID
  return(NWpropOverview)
}
```

## One network per dataset (BL and EA), kingdom (166S and 18S), and sizefraction (022_3 and 3_200)
### static network
```{r}
IDs = c("BL", "EA")
for(i in IDs)
{
  # data
  NW <- read.table(paste(PATH,"2_Networks/",i,"_temporalNetwork.tsv",sep=""), header = TRUE, stringsAsFactors = FALSE)
  G <- graph_from_edgelist(as.matrix(NW[,c("Source","Target")]))
  print(paste(i, ":",vcount(G), "nodes and", ecount(G), "edges"))
  
  # static network
  NWprop_static <- get_nw_proterties(NW, "Weight", "staticNW")
  NWprop_static <- t(NWprop_static)
  NWprop_static <- cbind(rownames(NWprop_static),NWprop_static)
  colnames(NWprop_static)[1] <- "Sample"

  # Save tables
  write.table(NWprop_static,
              paste(PATH,"3_NetworkProperties/NW_properties_overview_staticNW_",i,".tsv",sep=""),
              row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t", dec = ".")
}
```
BL : 1906 nodes and 2614 edges"
EA : 2074 nodes and 2389 edges"

Number of different types of edges in static network
```{r}
IDs = c("BL", "EA")
for(i in IDs)
{
  # data
  NW <- read.table(paste(PATH,"2_Networks/",i,"_temporalNetwork.tsv",sep=""), header = TRUE, stringsAsFactors = FALSE)[,c(1:2)]
  NW$Source <- gsub("OTU_[0-9]*_","", gsub("ASV_[0-9]*_","",NW$Source))
  NW$Target <- gsub("OTU_[0-9]*_","", gsub("ASV_[0-9]*_","",NW$Target))
  print(i)
  print(table(NW$Source, NW$Target))
}
```
[1] "BL"
            16S_022_3 16S_3_200 18S_022_3 18S_3_200
  16S_022_3       788       204       154       254
  16S_3_200         0        44         0        56
  18S_022_3         0        49       283       295
  18S_3_200         0         0         0       487
  
[1] "EA"
           
            16S_022_3 16S_3_200 18S_022_3 18S_3_200
  16S_022_3       217        31       180       127
  16S_3_200         0        14         0        37
  18S_022_3         0        51       695       614
  18S_3_200         0         0         0       423

### temporal network
```{r}
for(i in IDs)
{
  # data
  NW <- read.table(paste(PATH,"2_Networks/",i,"_temporalNetwork.tsv",sep=""), header = TRUE, stringsAsFactors = FALSE)
  G <- graph_from_edgelist(as.matrix(NW[,c("Source","Target")]))
  
  # temporal nw
  COLNAMES <- c(colnames(NW)[grep("BL",colnames(NW))], colnames(NW)[grep("EA",colnames(NW))])
  NWprop <- get_nw_proterties(NW[which(NW[,COLNAMES[1]]==1),], "Weight", COLNAMES[1])
  for(j in COLNAMES[-1]){
    if(dim(NW[which(NW[,j]==1),])[1]>1){
      NWpropnew <- get_nw_proterties(NW[which(NW[,j]==1),], "Weight", j)
      NWprop <- cbind(NWprop, NWpropnew)
    } else {
      #print(paste(i, "and", j, ": no edges", sep = ""))
      NWpropnew <- NA
      NWprop <- cbind(NWprop, NWpropnew)
      colnames(NWprop)[grep("NWpropnew",colnames(NWprop))] <- j
      NWprop["num_nodes",j] <- 0
      NWprop["num_edges",j] <- 0
    }
  }
  NWprop <- t(NWprop)
  NWprop <- cbind(rownames(NWprop),NWprop)
  colnames(NWprop)[1] <- "Sample"
  
  # Save tables
  write.table(NWprop,
              paste(PATH,"3_NetworkProperties/NW_properties_overview_temporalNW_",i,".tsv",sep=""),
              row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t", dec = ".")
}
```

Visualize: Network property change in time
- edge density
- average positive association score
- transitivity
- average path length
- assortativity (degree)
- assortativity (Bac-Euk)
```{r}
Month_labels <- c("01"="Jan", "02"="Feb", "03"="Mar", "04"="Apr", "05"="May", "06"="Jun",
                  "07"="Jul", "08"="Aug", "09"="Sep", "10"="Oct", "11"="Nov", "12"="Dec")
ALPHA=0.8
POINT_SIZE=2

for(i in IDs){
  NWprop <- read.table(paste(PATH,"3_NetworkProperties/NW_properties_overview_temporalNW_",i,".tsv",sep=""), header = TRUE)
  NWprop$Year <- substr(NWprop$Sample,6,7)
  NWprop$Month <- substr(NWprop$Sample,8,9)
  table(NWprop$Month)
  NWprop$Month_ID <- Month_labels[NWprop$Month]
  NWprop_stat <- read.table(paste(PATH,"3_NetworkProperties/NW_properties_overview_staticNW_",i,".tsv",sep=""), header = TRUE)

  dt_temp <- NWprop[,c("Sample", "Year", "Month", "num_nodes", "Month_ID")]
  p_temp <- ggplot(dt_temp, aes(x=factor(Sample, levels=rev(Sample)), y=num_nodes, label=Month_ID)) +
            geom_hline(yintercept = as.numeric(NWprop_stat$num_nodes), linetype = "dashed", colour = "black") +
            geom_point(color=Color_Months[dt_temp$Month_ID],size=POINT_SIZE,alpha=ALPHA) + 
            geom_text(size = 1) +
            ylab("edge density") + xlab("Samples") +
            coord_flip()
  ggsave(filename = paste(PATH,"3_NetworkProperties/",i,"/",i,"_num_nodes.pdf",sep=""), plot = p_temp, width = 4, height = 8)
  
  dt_temp <- NWprop[,c("Sample", "Year", "Month", "num_edges", "Month_ID")]
  p_temp <- ggplot(dt_temp, aes(x=factor(Sample, levels=rev(Sample)), y=num_edges, label=Month_ID)) +
            geom_hline(yintercept = as.numeric(NWprop_stat$num_edges), linetype = "dashed", colour = "black") +
            geom_point(color=Color_Months[dt_temp$Month_ID],size=POINT_SIZE,alpha=ALPHA) + 
            geom_text(size = 1) +
            ylab("edge density") + xlab("Samples") +
            coord_flip()
  ggsave(filename = paste(PATH,"3_NetworkProperties/",i,"/",i,"_num_edges.pdf",sep=""), plot = p_temp, width = 4, height = 8)

  dt_temp <- NWprop[,c("Sample", "Year", "Month", "edge_density", "Month_ID")]
  p_temp <- ggplot(dt_temp, aes(x=factor(Sample, levels=rev(Sample)), y=edge_density, label=Month_ID)) +
            geom_hline(yintercept = as.numeric(NWprop_stat$edge_density), linetype = "dashed", colour = "black") +
            geom_point(color=Color_Months[dt_temp$Month_ID],size=POINT_SIZE,alpha=ALPHA) + 
            geom_text(size = 1) +
            ylab("edge density") + xlab("Samples") +
            coord_flip()
  ggsave(filename = paste(PATH,"3_NetworkProperties/",i,"/",i,"_edge_density.pdf",sep=""), plot = p_temp, width = 4, height = 8)
  
  dt_temp <- NWprop[,c("Sample", "Year", "Month", "average_pos_association_strength", "Month_ID")]
  p_temp <- ggplot(dt_temp, aes(x=factor(Sample, levels=rev(Sample)), y=average_pos_association_strength, label=Year)) +
            geom_hline(yintercept = as.numeric(NWprop_stat$average_pos_association_strength), linetype = "dashed", colour = "black") +
            geom_point(color=Color_Months[dt_temp$Month_ID],size=POINT_SIZE,alpha=ALPHA) + 
            geom_text(size = 1) +
            ylab("avg pos ass strength") + xlab("Samples") +
            coord_flip()
  ggsave(filename = paste(PATH,"3_NetworkProperties/",i,"/",i,"_average_pos_association_strength.pdf",sep=""), plot = p_temp, width = 4, height = 8)
  
  dt_temp <- NWprop[,c("Sample", "Year", "Month", "transitivity_global_clusering_coef", "Month_ID")]
  p_temp <- ggplot(dt_temp, aes(x=factor(Sample, levels=rev(Sample)), y=transitivity_global_clusering_coef, label=Month_ID)) +
            geom_hline(yintercept = as.numeric(NWprop_stat$transitivity_global_clusering_coef), linetype = "dashed", colour = "black") +
            geom_point(color=Color_Months[dt_temp$Month_ID],size=POINT_SIZE,alpha=ALPHA) + 
            geom_text(size = 1) +
            ylab("transitivity") + xlab("Samples") +
            coord_flip()
  ggsave(filename = paste(PATH,"3_NetworkProperties/",i,"/",i,"_transitivity_global_clusering_coef.pdf",sep=""), plot = p_temp, width = 4, height = 8)
  
  dt_temp <- NWprop[,c("Sample", "Year", "Month", "ave_path_length", "Month_ID")]
  p_temp <- ggplot(dt_temp, aes(x=factor(Sample, levels=rev(Sample)), y=ave_path_length, label=Month_ID)) + 
            geom_hline(yintercept = as.numeric(NWprop_stat$ave_path_length), linetype = "dashed", colour = "black") +
            geom_point(color=Color_Months[dt_temp$Month_ID],size=POINT_SIZE,alpha=ALPHA) + 
            geom_text(size = 1) +
            ylab("avg path length") + xlab("Samples") +
            coord_flip()
  ggsave(filename = paste(PATH,"3_NetworkProperties/",i,"/",i,"_ave_path_length.pdf",sep=""), plot = p_temp, width = 4, height = 8)
  
  dt_temp <- NWprop[,c("Sample", "Year", "Month", "assortativity_degree", "Month_ID")]
  p_temp <- ggplot(dt_temp, aes(x=factor(Sample, levels=rev(Sample)), y=assortativity_degree, label=Month_ID)) + 
            geom_hline(yintercept = as.numeric(NWprop_stat$assortativity_degree), linetype = "dashed", colour = "black") +
            geom_point(color=Color_Months[dt_temp$Month_ID],size=POINT_SIZE,alpha=ALPHA) + 
            geom_text(size = 1.5) +
            ylab("assortativity (degree)") + xlab("Samples") +
            coord_flip()
  ggsave(filename = paste(PATH,"3_NetworkProperties/",i,"/",i,"_assortativity_degree.pdf",sep=""), plot = p_temp, width = 4, height = 8)
  
  dt_temp <- NWprop[,c("Sample", "Year", "Month", "assortativity_nominal_EukProk", "Month_ID")]
  p_temp <- ggplot(dt_temp, aes(x=factor(Sample, levels=rev(Sample)), y=assortativity_nominal_EukProk, label=Month_ID)) + 
            geom_hline(yintercept = as.numeric(NWprop_stat$assortativity_nominal_EukProk), linetype = "dashed", colour = "black") +
            geom_point(color=Color_Months[dt_temp$Month_ID],size=POINT_SIZE,alpha=ALPHA) + 
            geom_text(size = 1) +
            ylab("assortativity (Bac-Euk)") + xlab("Samples") +
            coord_flip()
  ggsave(filename = paste(PATH,"3_NetworkProperties/",i,"/",i,"_assortativity_nominal_EukProk.pdf",sep=""), plot = p_temp, width = 4, height = 8)
}
```
