---
title: "TaxonomyOfInteractions_Seasons"
author: "Ina Maria Deutschmann"
date: "06/02/2025"
output: html_document
---

# Taxonomy of interactions (Static Network)
## Set Up
Path and filenames
```{r}
PATH <- "~/Documents/Ina/Pedro/"

file_temporal_NW_BL <- "2_Networks/BL_temporalNetwork.tsv"
file_temporal_NW_EA <- "2_Networks/EA_temporalNetwork.tsv"
file_ASV16S <- "1_Data/ASV_16S_seqs_tax.tsv"
file_ASV18S <- "1_Data/ASV_18S_seqs_tax.tsv"
file_ENV <- "0_PedroData/EnvData_EAMO_BBMO_12112024.csv"
```

Load necessary libraries
```{r}
library(dplyr)
library(tidyr)
library(circlize)
library(RColorBrewer)
library(ggplot2)
library(knitr)
library(kableExtra)
library(flextable)
library(officer)
```

Load data
```{r}
# Load Networks
NW_BL <- read.table(paste0(PATH, file_temporal_NW_BL), header = TRUE, stringsAsFactors = FALSE)
NW_EA <- read.table(paste0(PATH, file_temporal_NW_EA), header = TRUE, stringsAsFactors = FALSE)

# Load Taxonomy Files
TAX_16S <- read.table(paste0(PATH, file_ASV16S), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
TAX_18S <- read.table(paste0(PATH, file_ASV18S), header = TRUE, sep = "\t", stringsAsFactors = FALSE)

# Set rownames for easy access
rownames(TAX_16S) <- TAX_16S$ID
rownames(TAX_18S) <- TAX_18S$ID

# Load Environmental file
ENV <- read.table(paste0(PATH, file_ENV), header = TRUE, sep = ",", stringsAsFactors = FALSE)

# Fix Sample IDs in ENV to match column names (replace '-' with '.')
ENV <- ENV %>%
  mutate(SampleID_fixed = gsub("-", ".", SampleID))
```

Test for missing values BL
```{r}
# Extract unique IDs from your network
network_ids <- unique(c(NW_BL$Source, NW_BL$Target))

# Clean IDs (remove the 16S/18S part to match TAX IDs)
network_ids_clean <- sub("_(16S_.*|18S_.*)", "", network_ids)

# Split by 16S and 18S
ids_16S <- network_ids_clean[grepl("16S", network_ids)]
ids_18S <- network_ids_clean[grepl("18S", network_ids)]

# Find missing IDs
missing_16S <- ids_16S[!ids_16S %in% rownames(TAX_16S)]
missing_18S <- ids_18S[!ids_18S %in% rownames(TAX_18S)]

# Combine and Print
missing_ids <- unique(c(missing_16S, missing_18S))

# Print missing IDs
cat("Missing IDs:\n")
print(missing_ids)

# Optional: How many missing
cat("\nNumber of missing IDs:", length(missing_ids), "\n")
```

Test for missing values EA
```{r}
# Extract unique IDs from your network
network_ids <- unique(c(NW_EA$Source, NW_EA$Target))

# Clean IDs (remove the 16S/18S part to match TAX IDs)
network_ids_clean <- sub("_(16S_.*|18S_.*)", "", network_ids)

# Split by 16S and 18S
ids_16S <- network_ids_clean[grepl("16S", network_ids)]
ids_18S <- network_ids_clean[grepl("18S", network_ids)]

# Find missing IDs
missing_16S <- ids_16S[!ids_16S %in% rownames(TAX_16S)]
missing_18S <- ids_18S[!ids_18S %in% rownames(TAX_18S)]

# Combine and Print
missing_ids <- unique(c(missing_16S, missing_18S))

# Print missing IDs
cat("Missing IDs:\n")
print(missing_ids)

# Optional: How many missing
cat("\nNumber of missing IDs:", length(missing_ids), "\n")
```
Test for NA in Phylum (or Class for Protobacteria) or Devision
```{r}
# 16S: Check for NAs in Phylum or Class
missing_phylum_16S <- TAX_16S %>%
  filter(is.na(Phylum) | (Phylum == "Proteobacteria" & is.na(Class))) %>%
  pull(ID)

# 18S: Check for NAs in Division
missing_division_18S <- TAX_18S %>%
  filter(is.na(Division)) %>%
  pull(ID)

# Combine
missing_taxonomy_ids <- unique(c(missing_phylum_16S, missing_division_18S))

# Print
cat("IDs with missing Phylum/Class/Division:\n")
print(missing_taxonomy_ids)

# How many
cat("\nNumber of IDs with missing taxonomy:", length(missing_taxonomy_ids), "\n")
```

Annotate Network with Taxonomy
```{r}
# Helper Function to Extract Taxonomy (Safe version with unknown labels)
get_division_phylum <- function(ID) {
  if (grepl("16S", ID)) {
    id_clean <- sub("_(16S_.*)", "", ID)
    if (id_clean %in% rownames(TAX_16S)) {
      tx <- TAX_16S[id_clean, ]
      if (is.na(tx$Phylum)) {
        division_phylum <- "unknown 16S"
      } else if (tx$Phylum == "Proteobacteria") {
        division_phylum <- ifelse(is.na(tx$Class), "unknown 16S", tx$Class)
      } else {
        division_phylum <- tx$Phylum
      }
      return(data.frame(Kingdom = tx$Kingdom, DivisionPhylum = division_phylum))
    } else {
      return(data.frame(Kingdom = NA, DivisionPhylum = "unknown 16S"))
    }
  } else if (grepl("18S", ID)) {
    id_clean <- sub("_(18S_.*)", "", ID)
    if (id_clean %in% rownames(TAX_18S)) {
      tx <- TAX_18S[id_clean, ]
      if (is.na(tx$Division)) {
        division_phylum <- "unknown 18S"
      } else {
        division_phylum <- tx$Division
      }
      return(data.frame(Kingdom = tx$Kingdom, DivisionPhylum = division_phylum))
    } else {
      return(data.frame(Kingdom = NA, DivisionPhylum = "unknown 18S"))
    }
  } else {
    return(data.frame(Kingdom = NA, DivisionPhylum = "unknown"))
  }
}

# Annotate Network with Taxonomy
NW_BL_annotated <- NW_BL %>%
  mutate(Source_ID = sub("_(16S_.*|18S_.*)", "", Source),
         Target_ID = sub("_(16S_.*|18S_.*)", "", Target)) %>%
  rowwise() %>%
  mutate(Source_Kingdom = get_division_phylum(Source)$Kingdom,
         Source_DivisionPhylum = get_division_phylum(Source)$DivisionPhylum,
         Target_Kingdom = get_division_phylum(Target)$Kingdom,
         Target_DivisionPhylum = get_division_phylum(Target)$DivisionPhylum) %>%
  ungroup()

NW_EA_annotated <- NW_EA %>%
  mutate(Source_ID = sub("_(16S_.*|18S_.*)", "", Source),
         Target_ID = sub("_(16S_.*|18S_.*)", "", Target)) %>%
  rowwise() %>%
  mutate(Source_Kingdom = get_division_phylum(Source)$Kingdom,
         Source_DivisionPhylum = get_division_phylum(Source)$DivisionPhylum,
         Target_Kingdom = get_division_phylum(Target)$Kingdom,
         Target_DivisionPhylum = get_division_phylum(Target)$DivisionPhylum) %>%
  ungroup()
```

```{r}
colnames(NW_BL_annotated)
```
```{r}
colnames(NW_EA_annotated)
```
```{r}
colnames(ENV)
```

## Map seasons
```{r}
# 1. Extract sample columns from NW_BL and NW_EA
sample_cols_BL <- colnames(NW_BL)[grep("^BL\\.", colnames(NW_BL))]
sample_cols_EA <- colnames(NW_EA)[grep("^EA\\.", colnames(NW_EA))]

# 2. Combine all sample names
all_samples_NW <- c(sample_cols_BL, sample_cols_EA)

# 3. Subset ENV for those samples
env_sub <- ENV %>%
  filter(SampleID %in% all_samples_NW_fixed)

# 4. Summarize number of samples per station and season
season_site_summary <- env_sub %>%
  group_by(station, season) %>%
  summarise(n_samples = n()) %>%
  arrange(station, season)

# 5. Show the result
kable(season_site_summary, caption = "Number of Samples per Season and Site (Only for Network Samples)") %>%
  kable_styling(full_width = FALSE)
```

Number of Samples per Season and Site (Only for Network Samples)
station	season	n_samples
Blanes Bay Microbial Observatory	Autumn	6
Blanes Bay Microbial Observatory	Spring	5
Blanes Bay Microbial Observatory	Summer	4
Blanes Bay Microbial Observatory	Winter	8
Equatorial Atlantic Microbial Observatory	Autumn	8
Equatorial Atlantic Microbial Observatory	Spring	5
Equatorial Atlantic Microbial Observatory	Summer	3
Equatorial Atlantic Microbial Observatory	Winter	6

## Merge subnetworks per season and site
```{r}
# Function to merge subnetworks by season for a given site
merge_network_by_season <- function(NW, ENV, site_prefix, season_name) {
  # 1. Get samples for this site and season
  samples_season <- ENV %>%
    filter(grepl(site_prefix, SampleID), season == season_name) %>%
    pull(SampleID_fixed)
  
  # 2. Keep only samples that exist in the network columns
  samples_season_exist <- samples_season[samples_season %in% colnames(NW)]
  
  # 3. Columns to keep: Source, Target, and samples of that season
  columns_to_keep <- c("Source", "Target", samples_season_exist,
                       "Source_Kingdom", "Source_DivisionPhylum",
                       "Target_Kingdom","Target_DivisionPhylum")
  
  # 4. Subset the network
  NW_season <- NW[, columns_to_keep, drop = FALSE]
  
  # 5. Calculate occurrence
  if (length(samples_season_exist) > 0) {
    NW_season$Occurrence <- rowSums(NW_season[, samples_season_exist] > 0)
  } else {
    NW_season$Occurrence <- 0
  }

  # 6. Remove edges with 0 occurrence
  NW_season <- NW_season %>%
    filter(Occurrence > 0)
  
  # 7. Add Occurrence Percentage (rounded to 2 decimal digits)
  NW_season$Occurrence_Percentage <- round((NW_season$Occurrence / length(samples_season_exist)) * 100, 2)
  
  # 8. Return
  NW_season_result <- NW_season[, c("Source", "Target", "Occurrence", "Occurrence_Percentage",
                                    "Source_Kingdom", "Source_DivisionPhylum",
                                    "Target_Kingdom","Target_DivisionPhylum")]
  
  return(NW_season_result)
}

# Merge networks
NW_BL_Winter <- merge_network_by_season(NW_BL_annotated, ENV, "BL", "Winter")
NW_BL_Spring <- merge_network_by_season(NW_BL_annotated, ENV, "BL", "Spring")
NW_BL_Summer <- merge_network_by_season(NW_BL_annotated, ENV, "BL", "Summer")
NW_BL_Autumn <- merge_network_by_season(NW_BL_annotated, ENV, "BL", "Autumn")

NW_EA_Winter <- merge_network_by_season(NW_EA_annotated, ENV, "EA", "Winter")
NW_EA_Spring <- merge_network_by_season(NW_EA_annotated, ENV, "EA", "Spring")
NW_EA_Summer <- merge_network_by_season(NW_EA_annotated, ENV, "EA", "Summer")
NW_EA_Autumn <- merge_network_by_season(NW_EA_annotated, ENV, "EA", "Autumn")
```

```{r}
head(NW_EA_Summer)
```

Check merging on an example
```{r}
# 1. Define Summer samples for EA
summer_samples_EA <- ENV %>%
  filter(grepl("^EA", SampleID), season == "Summer") %>%
  pull(SampleID_fixed)

# 2. Check which samples exist in NW_EA
summer_samples_EA_exist <- summer_samples_EA[summer_samples_EA %in% colnames(NW_EA)]

# 3. Columns to select
columns_to_select <- c("Source", "Target", summer_samples_EA_exist)

# 4. Subset
NW_EA_summer_subset <- NW_EA[, columns_to_select, drop = FALSE]

# 5. Print first 5 rows for comparison
cat("\n--- Raw Summer Subset (First 5 rows) ---\n")
print(head(NW_EA_summer_subset, 5))

cat("\n--- Merged Summer Network (First 5 rows) ---\n")
print(head(NW_EA_Summer, 5))
```
Correct

## Overview occurrence
```{r}
# Add site and season labels
add_labels <- function(df, site, season) {
  df %>%
    mutate(Site = site, Season = season)
}

# Combine all season + site datasets
NW_all <- bind_rows(
  add_labels(NW_BL_Winter, "BL", "Winter"),
  add_labels(NW_BL_Spring, "BL", "Spring"),
  add_labels(NW_BL_Summer, "BL", "Summer"),
  add_labels(NW_BL_Autumn, "BL", "Autumn"),
  add_labels(NW_EA_Winter, "EA", "Winter"),
  add_labels(NW_EA_Spring, "EA", "Spring"),
  add_labels(NW_EA_Summer, "EA", "Summer"),
  add_labels(NW_EA_Autumn, "EA", "Autumn")
)

# Plot
p <- ggplot(NW_all, aes(x = Occurrence_Percentage)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "black") +  # 5% bins
  facet_grid(Site ~ Season) +
  xlab("Occurrence (%)") +
  ylab("Number of Edges") +
  theme_bw() +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  ) +
  ggtitle("Distribution of Edge Occurrence Percentage by Season and Site")

# Show plot in RStudio Viewer
print(p)

# Save to PDF
pdf(file = paste0(PATH, "/5_NetworkVisualization/AssociationOccurrence_Histogram_by_Season_and_Site.pdf"),
    width = 10, height = 6)
print(p)
dev.off()
```

## Save seasonal networks with taxonomy and occurence
```{r}
# Create a helper function to save a network
save_network <- function(df, site, season) {
  filename <- paste0(PATH, "2_Networks/", site, "_", season, "_SeasonalNetwork_Taxonomy.tsv")
  write.table(df, file = filename, sep = "\t", quote = FALSE, row.names = FALSE)
}

# Save all 8 networks
save_network(NW_BL_Winter, "BL", "Winter")
save_network(NW_BL_Spring, "BL", "Spring")
save_network(NW_BL_Summer, "BL", "Summer")
save_network(NW_BL_Autumn, "BL", "Autumn")

save_network(NW_EA_Winter, "EA", "Winter")
save_network(NW_EA_Spring, "EA", "Spring")
save_network(NW_EA_Summer, "EA", "Summer")
save_network(NW_EA_Autumn, "EA", "Autumn")
```

## Chord plot
Filter based on occurrence_percentage > 75%
```{r}
# Helper function to filter a network based on Occurrence_Percentage
filter_high_occurrence <- function(df, threshold = 75) {
  df %>%
    filter(Occurrence_Percentage > threshold)
}

# Apply to all
NW_BL_Winter_filt <- filter_high_occurrence(NW_BL_Winter)
NW_BL_Spring_filt <- filter_high_occurrence(NW_BL_Spring)
NW_BL_Summer_filt <- filter_high_occurrence(NW_BL_Summer)
NW_BL_Autumn_filt <- filter_high_occurrence(NW_BL_Autumn)

NW_EA_Winter_filt <- filter_high_occurrence(NW_EA_Winter)
NW_EA_Spring_filt <- filter_high_occurrence(NW_EA_Spring)
NW_EA_Summer_filt <- filter_high_occurrence(NW_EA_Summer)
NW_EA_Autumn_filt <- filter_high_occurrence(NW_EA_Autumn)
```

Generate interaction tables
```{r}
# Generate interaction table for chord diagram
prepare_interaction_table <- function(df) {
  df %>%
    count(Source_DivisionPhylum, Target_DivisionPhylum) %>%
    rename(Weight = n) %>%
    rowwise() %>%
    mutate(
      Source_sorted = min(Source_DivisionPhylum, Target_DivisionPhylum),
      Target_sorted = max(Source_DivisionPhylum, Target_DivisionPhylum)
    ) %>%
    group_by(Source_sorted, Target_sorted) %>%
    summarise(Weight = sum(Weight), .groups = 'drop')
}

interaction_BL_Winter <- prepare_interaction_table(NW_BL_Winter_filt)
interaction_BL_Spring <- prepare_interaction_table(NW_BL_Spring_filt)
interaction_BL_Summer <- prepare_interaction_table(NW_BL_Summer_filt)
interaction_BL_Autumn <- prepare_interaction_table(NW_BL_Autumn_filt)

interaction_EA_Winter <- prepare_interaction_table(NW_EA_Winter_filt)
interaction_EA_Spring <- prepare_interaction_table(NW_EA_Spring_filt)
interaction_EA_Summer <- prepare_interaction_table(NW_EA_Summer_filt)
interaction_EA_Autumn <- prepare_interaction_table(NW_EA_Autumn_filt)
```

Store interaction tables
```{r}
write.table(interaction_BL_Winter, 
            file = paste0(PATH, "/2_Networks/SeasonNW_Above75percentOccurrence_BL_InteractionTable_Winter.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_BL_Spring, 
            file = paste0(PATH, "/2_Networks/SeasonNW_Above75percentOccurrence_BL_InteractionTable_Spring.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_BL_Summer, 
            file = paste0(PATH, "/2_Networks/SeasonNW_Above75percentOccurrence_BL_InteractionTable_Summer.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_BL_Autumn, 
            file = paste0(PATH, "/2_Networks/SeasonNW_Above75percentOccurrence_BL_InteractionTable_Autumn.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_EA_Winter, 
            file = paste0(PATH, "/2_Networks/SeasonNW_Above75percentOccurrence_EA_InteractionTable_Winter.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_EA_Spring, 
            file = paste0(PATH, "/2_Networks/SeasonNW_Above75percentOccurrence_EA_InteractionTable_Spring.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_EA_Summer, 
            file = paste0(PATH, "/2_Networks/SeasonNW_Above75percentOccurrence_EA_InteractionTable_Summer.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_EA_Autumn, 
            file = paste0(PATH, "/2_Networks/SeasonNW_Above75percentOccurrence_EA_InteractionTable_Autumn.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)
```

Shared color palette
```{r}
# Extract all DivisionPhylum from filtered seasonal networks
Groups_all_filtered <- sort(unique(c(
  NW_BL_Winter_filt$Source_DivisionPhylum, NW_BL_Winter_filt$Target_DivisionPhylum,
  NW_BL_Spring_filt$Source_DivisionPhylum, NW_BL_Spring_filt$Target_DivisionPhylum,
  NW_BL_Summer_filt$Source_DivisionPhylum, NW_BL_Summer_filt$Target_DivisionPhylum,
  NW_BL_Autumn_filt$Source_DivisionPhylum, NW_BL_Autumn_filt$Target_DivisionPhylum,
  NW_EA_Winter_filt$Source_DivisionPhylum, NW_EA_Winter_filt$Target_DivisionPhylum,
  NW_EA_Spring_filt$Source_DivisionPhylum, NW_EA_Spring_filt$Target_DivisionPhylum,
  NW_EA_Summer_filt$Source_DivisionPhylum, NW_EA_Summer_filt$Target_DivisionPhylum,
  NW_EA_Autumn_filt$Source_DivisionPhylum, NW_EA_Autumn_filt$Target_DivisionPhylum
)))

# Color palette
colors_all_filtered <- colorRampPalette(brewer.pal(8, "Set2"))(length(Groups_all_filtered))
grid_colors_all_filtered <- setNames(colors_all_filtered, Groups_all_filtered)
```

Label abbreviation
```{r}
# Step 1: Abbreviations: first 3 letters
abbreviations_filtered <- substr(Groups_all_filtered, 1, 3)

# Step 2: Create a table with full name and abbreviation
abbr_table_filtered <- data.frame(
  FullName = Groups_all_filtered,
  Abbreviation = abbreviations_filtered,
  stringsAsFactors = FALSE
)

# Step 3: Check for abbreviations linked to multiple full names
duplicate_abbr_filtered <- abbr_table_filtered %>%
  group_by(Abbreviation) %>%
  summarise(UniqueFullNames = n_distinct(FullName)) %>%
  filter(UniqueFullNames > 1)

# Step 4: List only problematic abbreviations
problematic_abbr_filtered <- duplicate_abbr_filtered$Abbreviation

# Step 5: Show problematic entries
conflicts_filtered <- abbr_table_filtered %>%
  filter(Abbreviation %in% problematic_abbr_filtered) %>%
  arrange(Abbreviation)

cat("Non-unique abbreviations (different long names):\n")
print(conflicts_filtered)
```
There are conflicts:
- Chloroflexi	Chl
- Chlorophyta	Chl

Custom abbreviation
```{r}
# Step 6: Custom abbreviations for conflicts (only for new conflicts found)
custom_abbr_filtered <- c(
  "Chloroflexi"  = "Chlf",  # 4 letters
  "Chlorophyta"  = "Chlp",   # 4 letters
  "Rhodelphidia"  = "Rdi",   # 3 letters
  "Rhodophyta"    = "Rph",   # 3 letters
  "unknown 16S"   = "u16",   # 3 letters
  "unknown 18S"   = "u18"    # 3 letters
  # Add more if any other conflicts appear
)

# Step 7: Create the abbreviation map
# Start: default (first 3 letters)
abbreviation_map_filtered <- setNames(substr(Groups_all_filtered, 1, 3), Groups_all_filtered)

# Step 8: Overwrite conflicts with custom ones
abbreviation_map_filtered[names(custom_abbr_filtered)] <- custom_abbr_filtered
```

Chord Diagrams: one per season and site
```{r}
# Function to plot seasonal chord diagrams
plot_season_chord <- function(interaction_table, site, season, PATH) {
  
  # Set PDF file name
  pdf_file <- paste0(PATH, "/5_NetworkVisualization/ChordDiagram_Seasons_", site, "_", season, ".pdf")
  
  # Open PDF
  pdf(file = pdf_file, width = 8, height = 8)
  
  # Clear previous plot
  circos.clear()
  
  # Set circos parameters
  circos.par(
    gap.after = rep(5, length(unique(c(interaction_table$Source_sorted, interaction_table$Target_sorted)))),
    start.degree = 90,   # Start at the top
    track.margin = c(0.05, 0.05),
    cell.padding = c(0, 0, 0, 0)
  )
  
  # Draw chord diagram
  chordDiagram(
    interaction_table %>% select(Source_sorted, Target_sorted, Weight),
    grid.col = grid_colors_all_filtered,
    transparency = 0.3,
    annotationTrack = "grid"
  )
  
  # Add labels (abbreviations)
  circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
    sector.name <- get.cell.meta.data("sector.index")
    circos.text(
      x = mean(get.cell.meta.data("xlim")),
      y = 1.1,
      labels = abbreviation_map_filtered[sector.name],
      facing = "clockwise",
      niceFacing = TRUE,
      adj = c(0, 0.5),
      cex = 0.7
    )
  }, bg.border = NA)
  
  # Close PDF
  dev.off()
  
  # Clear circos plot
  circos.clear()
  
  message(paste("Saved:", pdf_file))
}
```

```{r}
plot_season_chord(interaction_BL_Winter, site = "BL", season = "Winter", PATH = PATH)
plot_season_chord(interaction_BL_Spring, site = "BL", season = "Spring", PATH = PATH)
plot_season_chord(interaction_BL_Summer, site = "BL", season = "Summer", PATH = PATH)
plot_season_chord(interaction_BL_Autumn, site = "BL", season = "Autumn", PATH = PATH)

plot_season_chord(interaction_EA_Winter, site = "EA", season = "Winter", PATH = PATH)
plot_season_chord(interaction_EA_Spring, site = "EA", season = "Spring", PATH = PATH)
plot_season_chord(interaction_EA_Summer, site = "EA", season = "Summer", PATH = PATH)
plot_season_chord(interaction_EA_Autumn, site = "EA", season = "Autumn", PATH = PATH)
```

Legend
```{r}
# Create a DataFrame for the legend (Filtered Version)
legend_df_filtered <- data.frame(
  Color = grid_colors_all_filtered[Groups_all_filtered],            # HEX color codes
  Abbreviation = abbreviation_map_filtered[Groups_all_filtered],    # 3-letter (or custom) abbreviation
  FullName = Groups_all_filtered,                                   # Full taxonomic name
  stringsAsFactors = FALSE
)

# Save as CSV
write.csv(legend_df_filtered, file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_Seasons_Legend.csv"), row.names = FALSE)

# Display simple kable
kable(legend_df_filtered)
```

Color swatch legend
```{r}
# Color swatch kable
legend_df_filtered %>%
  mutate(Color = cell_spec(" ", background = Color)) %>%
  kable(escape = FALSE, align = "c", format = "html") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))
```

Save as WORD
```{r}
# Create flextable
ft_filtered <- flextable(legend_df_filtered)

# Set background color for the Color column
for (i in seq_len(nrow(legend_df_filtered))) {
  ft_filtered <- bg(ft_filtered, i = i, j = 1, bg = legend_df_filtered$Color[i])
}

# Set text color for Color column (white text for better contrast)
ft_filtered <- color(ft_filtered, j = 1, color = "white")

# Autofit
ft_filtered <- autofit(ft_filtered)

# Show in Viewer
ft_filtered

# Save as Word
doc_filtered <- read_docx()
doc_filtered <- body_add_flextable(doc_filtered, ft_filtered)
print(doc_filtered, target = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_Seasons_Legend.docx"))
```
