---
title: "TaxonomyOfInteractions"
author: "Ina Maria Deutschmann"
date: "06/01/2025"
output: html_document
---

# Taxonomy of interactions (Static Network)
Path and filenames
```{r}
PATH <- "~/Documents/Ina/Pedro/"

file_static_NW_BL <- "2_Networks/BL_NW_static.tsv"
file_static_NW_EA <- "2_Networks/EA_NW_static.tsv"
file_ASV16S <- "1_Data/ASV_16S_seqs_tax.tsv"
file_ASV18S <- "1_Data/ASV_18S_seqs_tax.tsv"
```

Load necessary libraries
```{r}
library(dplyr)
library(tidyr)
library(circlize)
library(RColorBrewer)
library(knitr)
library(kableExtra)
library(flextable)
library(officer)
```

Load data
```{r}
# Load Networks
NW_BL <- read.table(paste0(PATH, file_static_NW_BL), header = TRUE, stringsAsFactors = FALSE)
NW_EA <- read.table(paste0(PATH, file_static_NW_EA), header = TRUE, stringsAsFactors = FALSE)

# Load Taxonomy Files
TAX_16S <- read.table(paste0(PATH, file_ASV16S), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
TAX_18S <- read.table(paste0(PATH, file_ASV18S), header = TRUE, sep = "\t", stringsAsFactors = FALSE)

# Set rownames for easy access
rownames(TAX_16S) <- TAX_16S$ID
rownames(TAX_18S) <- TAX_18S$ID
```

Test for missing values BL
```{r}
# Extract unique IDs from your network
network_ids <- unique(c(NW_BL$Source, NW_BL$Target))

# Clean IDs (remove the 16S/18S part to match TAX IDs)
network_ids_clean <- sub("_(16S_.*|18S_.*)", "", network_ids)

# Split by 16S and 18S
ids_16S <- network_ids_clean[grepl("16S", network_ids)]
ids_18S <- network_ids_clean[grepl("18S", network_ids)]

# Find missing IDs
missing_16S <- ids_16S[!ids_16S %in% rownames(TAX_16S)]
missing_18S <- ids_18S[!ids_18S %in% rownames(TAX_18S)]

# Combine and Print
missing_ids <- unique(c(missing_16S, missing_18S))

# Print missing IDs
cat("Missing IDs:\n")
print(missing_ids)

# Optional: How many missing
cat("\nNumber of missing IDs:", length(missing_ids), "\n")
```

Test for missing values EA
```{r}
# Extract unique IDs from your network
network_ids <- unique(c(NW_EA$Source, NW_EA$Target))

# Clean IDs (remove the 16S/18S part to match TAX IDs)
network_ids_clean <- sub("_(16S_.*|18S_.*)", "", network_ids)

# Split by 16S and 18S
ids_16S <- network_ids_clean[grepl("16S", network_ids)]
ids_18S <- network_ids_clean[grepl("18S", network_ids)]

# Find missing IDs
missing_16S <- ids_16S[!ids_16S %in% rownames(TAX_16S)]
missing_18S <- ids_18S[!ids_18S %in% rownames(TAX_18S)]

# Combine and Print
missing_ids <- unique(c(missing_16S, missing_18S))

# Print missing IDs
cat("Missing IDs:\n")
print(missing_ids)

# Optional: How many missing
cat("\nNumber of missing IDs:", length(missing_ids), "\n")
```
Test for NA in Phylum (or Class for Protobacteria) or Devision
```{r}
# 16S: Check for NAs in Phylum or Class
missing_phylum_16S <- TAX_16S %>%
  filter(is.na(Phylum) | (Phylum == "Proteobacteria" & is.na(Class))) %>%
  pull(ID)

# 18S: Check for NAs in Division
missing_division_18S <- TAX_18S %>%
  filter(is.na(Division)) %>%
  pull(ID)

# Combine
missing_taxonomy_ids <- unique(c(missing_phylum_16S, missing_division_18S))

# Print
cat("IDs with missing Phylum/Class/Division:\n")
print(missing_taxonomy_ids)

# How many
cat("\nNumber of IDs with missing taxonomy:", length(missing_taxonomy_ids), "\n")
```

Annotate Network with Taxonomy
```{r}
# Helper Function to Extract Taxonomy (Safe version with unknown labels)
get_division_phylum <- function(ID) {
  if (grepl("16S", ID)) {
    id_clean <- sub("_(16S_.*)", "", ID)
    if (id_clean %in% rownames(TAX_16S)) {
      tx <- TAX_16S[id_clean, ]
      if (is.na(tx$Phylum)) {
        division_phylum <- "unknown 16S"
      } else if (tx$Phylum == "Proteobacteria") {
        division_phylum <- ifelse(is.na(tx$Class), "unknown 16S", tx$Class)
      } else {
        division_phylum <- tx$Phylum
      }
      return(data.frame(Kingdom = tx$Kingdom, DivisionPhylum = division_phylum))
    } else {
      return(data.frame(Kingdom = NA, DivisionPhylum = "unknown 16S"))
    }
  } else if (grepl("18S", ID)) {
    id_clean <- sub("_(18S_.*)", "", ID)
    if (id_clean %in% rownames(TAX_18S)) {
      tx <- TAX_18S[id_clean, ]
      if (is.na(tx$Division)) {
        division_phylum <- "unknown 18S"
      } else {
        division_phylum <- tx$Division
      }
      return(data.frame(Kingdom = tx$Kingdom, DivisionPhylum = division_phylum))
    } else {
      return(data.frame(Kingdom = NA, DivisionPhylum = "unknown 18S"))
    }
  } else {
    return(data.frame(Kingdom = NA, DivisionPhylum = "unknown"))
  }
}

# Annotate Network with Taxonomy
NW_BL_annotated <- NW_BL %>%
  mutate(Source_ID = sub("_(16S_.*|18S_.*)", "", Source),
         Target_ID = sub("_(16S_.*|18S_.*)", "", Target)) %>%
  rowwise() %>%
  mutate(Source_Kingdom = get_division_phylum(Source)$Kingdom,
         Source_DivisionPhylum = get_division_phylum(Source)$DivisionPhylum,
         Target_Kingdom = get_division_phylum(Target)$Kingdom,
         Target_DivisionPhylum = get_division_phylum(Target)$DivisionPhylum) %>%
  ungroup()

NW_EA_annotated <- NW_EA %>%
  mutate(Source_ID = sub("_(16S_.*|18S_.*)", "", Source),
         Target_ID = sub("_(16S_.*|18S_.*)", "", Target)) %>%
  rowwise() %>%
  mutate(Source_Kingdom = get_division_phylum(Source)$Kingdom,
         Source_DivisionPhylum = get_division_phylum(Source)$DivisionPhylum,
         Target_Kingdom = get_division_phylum(Target)$Kingdom,
         Target_DivisionPhylum = get_division_phylum(Target)$DivisionPhylum) %>%
  ungroup()
```

Split based on interaction type
```{r}
NW_BL_annotated <- NW_BL_annotated %>%
  mutate(InteractionType = case_when(
    Source_Kingdom == "Bacteria" & Target_Kingdom == "Bacteria" ~ "Prokaryote-Prokaryote",
    Source_Kingdom == "Eukaryota" & Target_Kingdom == "Eukaryota" ~ "Eukaryote-Eukaryote",
    TRUE ~ "Cross-domain"
  ))

NW_prok_prok_BL <- NW_BL_annotated %>% filter(InteractionType == "Prokaryote-Prokaryote")
NW_euk_euk_BL   <- NW_BL_annotated %>% filter(InteractionType == "Eukaryote-Eukaryote")
NW_cross_BL     <- NW_BL_annotated %>% filter(InteractionType == "Cross-domain")

NW_EA_annotated <- NW_EA_annotated %>%
  mutate(InteractionType = case_when(
    Source_Kingdom == "Bacteria" & Target_Kingdom == "Bacteria" ~ "Prokaryote-Prokaryote",
    Source_Kingdom == "Eukaryota" & Target_Kingdom == "Eukaryota" ~ "Eukaryote-Eukaryote",
    TRUE ~ "Cross-domain"
  ))

NW_prok_prok_EA <- NW_EA_annotated %>% filter(InteractionType == "Prokaryote-Prokaryote")
NW_euk_euk_EA   <- NW_EA_annotated %>% filter(InteractionType == "Eukaryote-Eukaryote")
NW_cross_EA     <- NW_EA_annotated %>% filter(InteractionType == "Cross-domain")
```

Split interaction tables
```{r}
# Prokaryote-Prokaryote BL
interaction_table_prok_prok_BL <- NW_prok_prok_BL %>%
  count(Source_DivisionPhylum, Target_DivisionPhylum) %>%
  rename(Weight = n) %>%
  rowwise() %>%
  mutate(Source_sorted = min(Source_DivisionPhylum, Target_DivisionPhylum),
         Target_sorted = max(Source_DivisionPhylum, Target_DivisionPhylum)) %>%
  group_by(Source_sorted, Target_sorted) %>%
  summarise(Weight = sum(Weight)) %>%
  ungroup()

# Eukaryote-Eukaryote BL
interaction_table_euk_euk_BL <- NW_euk_euk_BL %>%
  count(Source_DivisionPhylum, Target_DivisionPhylum) %>%
  rename(Weight = n) %>%
  rowwise() %>%
  mutate(Source_sorted = min(Source_DivisionPhylum, Target_DivisionPhylum),
         Target_sorted = max(Source_DivisionPhylum, Target_DivisionPhylum)) %>%
  group_by(Source_sorted, Target_sorted) %>%
  summarise(Weight = sum(Weight)) %>%
  ungroup()

# Cross-domain BL
interaction_table_cross_BL <- NW_cross_BL %>%
  count(Source_DivisionPhylum, Target_DivisionPhylum) %>%
  rename(Weight = n) %>%
  rowwise() %>%
  mutate(Source_sorted = min(Source_DivisionPhylum, Target_DivisionPhylum),
         Target_sorted = max(Source_DivisionPhylum, Target_DivisionPhylum)) %>%
  group_by(Source_sorted, Target_sorted) %>%
  summarise(Weight = sum(Weight)) %>%
  ungroup()
```

```{r}
# Prokaryote-Prokaryote EA
interaction_table_prok_prok_EA <- NW_prok_prok_EA %>%
  count(Source_DivisionPhylum, Target_DivisionPhylum) %>%
  rename(Weight = n) %>%
  rowwise() %>%
  mutate(Source_sorted = min(Source_DivisionPhylum, Target_DivisionPhylum),
         Target_sorted = max(Source_DivisionPhylum, Target_DivisionPhylum)) %>%
  group_by(Source_sorted, Target_sorted) %>%
  summarise(Weight = sum(Weight)) %>%
  ungroup()

# Eukaryote-Eukaryote EA
interaction_table_euk_euk_EA <- NW_euk_euk_EA %>%
  count(Source_DivisionPhylum, Target_DivisionPhylum) %>%
  rename(Weight = n) %>%
  rowwise() %>%
  mutate(Source_sorted = min(Source_DivisionPhylum, Target_DivisionPhylum),
         Target_sorted = max(Source_DivisionPhylum, Target_DivisionPhylum)) %>%
  group_by(Source_sorted, Target_sorted) %>%
  summarise(Weight = sum(Weight)) %>%
  ungroup()

# Cross-domain EA
interaction_table_cross_EA <- NW_cross_EA %>%
  count(Source_DivisionPhylum, Target_DivisionPhylum) %>%
  rename(Weight = n) %>%
  rowwise() %>%
  mutate(Source_sorted = min(Source_DivisionPhylum, Target_DivisionPhylum),
         Target_sorted = max(Source_DivisionPhylum, Target_DivisionPhylum)) %>%
  group_by(Source_sorted, Target_sorted) %>%
  summarise(Weight = sum(Weight)) %>%
  ungroup()
```

Store interaction tables
```{r}
write.table(interaction_table_prok_prok_BL, 
            file = paste0(PATH, "/2_Networks/StaticNW_BL_InteractionTable_ProkProk.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_table_euk_euk_BL, 
            file = paste0(PATH, "/2_Networks/StaticNW_BLInteractionTable_EukEuk.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_table_cross_BL, 
            file = paste0(PATH, "/2_Networks/StaticNW_BLInteractionTable_Cross.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_table_prok_prok_EA, 
            file = paste0(PATH, "/2_Networks/StaticNW_EA_InteractionTable_ProkProk.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_table_euk_euk_EA, 
            file = paste0(PATH, "/2_Networks/StaticNW_EA_InteractionTable_EukEuk.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)

write.table(interaction_table_cross_EA, 
            file = paste0(PATH, "/2_Networks/StaticNW_EA_InteractionTable_Cross.tsv"),
            sep = "\t", row.names = FALSE, quote = FALSE)
```

Shared color palette
```{r}
Groups_BL_all <- unique(c(NW_BL_annotated$Source_DivisionPhylum, NW_BL_annotated$Target_DivisionPhylum))
Groups_EA_all <- unique(c(NW_EA_annotated$Source_DivisionPhylum, NW_EA_annotated$Target_DivisionPhylum))
Groups_all <- sort(unique(c(Groups_BL_all, Groups_EA_all)))

# Color palette
colors_all <- colorRampPalette(brewer.pal(8, "Set2"))(length(Groups_all))
grid_colors_all <- setNames(colors_all, Groups_all)
```

Label abbreviation
```{r}
# Abbreviations: first 3 letters
abbreviations <- substr(Groups_all, 1, 3)

# Step 2: Create a table with full name and abbreviation
abbr_table <- data.frame(
  FullName = Groups_all,
  Abbreviation = abbreviations,
  stringsAsFactors = FALSE
)

# Step 3: Check for abbreviations linked to multiple full names
# -> group by abbreviation and count how many distinct full names exist
duplicate_abbr <- abbr_table %>%
  group_by(Abbreviation) %>%
  summarise(UniqueFullNames = n_distinct(FullName)) %>%
  filter(UniqueFullNames > 1)

# Step 4: List only problematic abbreviations
problematic_abbr <- duplicate_abbr$Abbreviation

# Step 5: Show problematic entries
conflicts <- abbr_table %>%
  filter(Abbreviation %in% problematic_abbr) %>%
  arrange(Abbreviation)

cat("Non-unique abbreviations (different long names):")
print(conflicts)
```
There are conflicts:
- Chloroflexi	Chl			
- Chlorophyta	Chl			
- Rhodelphidia	Rho			
- Rhodophyta	Rho			
- unknown 16S	unk			
- unknown 18S	unk

Custom abbreviation
```{r}
# Custom abbreviations for conflicts
custom_abbr <- c(
  "Chloroflexi"   = "Chlf",  # 4 letters
  "Chlorophyta"   = "Chlp",  # 4 letters
  "Rhodelphidia"  = "Rdi",   # 3 letters
  "Rhodophyta"    = "Rph",   # 3 letters
  "unknown 16S"   = "u16",   # 3 letters
  "unknown 18S"   = "u18"    # 3 letters
  # Add more if any other conflicts appear
)

# Start: default abbreviations
abbreviation_map <- setNames(substr(Groups_all, 1, 3), Groups_all)

# Overwrite conflicts with custom ones
abbreviation_map[names(custom_abbr)] <- custom_abbr
```

Chord Diagram: Prokaryote-Prokaryote BL
```{r}
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_prok_prok_BL$Source_sorted, interaction_table_prok_prok_BL$Target_sorted)))),
  start.degree = 90,  # 90 degrees moves 0° to the top
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0))
chordDiagram(
  interaction_table_prok_prok_BL %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_BL_ProkProk_all.pdf"))
chordDiagram(
  interaction_table_prok_prok_BL %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```
Only show edges with weight ≥ 3
```{r}
# Filter: Keep only interactions with weight >= 3
interaction_table_filtered <- interaction_table_prok_prok_BL %>%
  filter(Weight >= 3)

# Create Chord Diagram
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_filtered$Source_sorted, interaction_table_filtered$Target_sorted)))),
  start.degree = 90,
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0)
)
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_BL_ProkProk_atLeast3.pdf"))
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```

Chord Diagram: Eukaryote-Eukaryote BL
```{r}
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_euk_euk_BL$Source_sorted, interaction_table_euk_euk_BL$Target_sorted)))),
  start.degree = 90,  # 90 degrees moves 0° to the top
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0))
chordDiagram(
  interaction_table_euk_euk_BL %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_BL_EukEuk_all.pdf"))
chordDiagram(
  interaction_table_euk_euk_BL %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```

Only show edges with weight ≥ 3
```{r}
# Filter: Keep only interactions with weight >= 3
interaction_table_filtered <- interaction_table_euk_euk_BL %>%
  filter(Weight >= 3)

# Create Chord Diagram
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_filtered$Source_sorted, interaction_table_filtered$Target_sorted)))),
  start.degree = 90,
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0)
)
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_BL_EukEuk_atLeast3.pdf"))
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```

Chord Diagram: Prokaryote-Eukaryote BL
```{r}
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_cross_BL$Source_sorted, interaction_table_cross_BL$Target_sorted)))),
  start.degree = 90,  # 90 degrees moves 0° to the top
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0))
chordDiagram(
  interaction_table_cross_BL %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_BL_EukProk_all.pdf"))
chordDiagram(
  interaction_table_cross_BL %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```
Only show edges with weight ≥ 3
```{r}
# Filter: Keep only interactions with weight >= 3
interaction_table_filtered <- interaction_table_cross_BL %>%
  filter(Weight >= 3)

# Create Chord Diagram
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_filtered$Source_sorted, interaction_table_filtered$Target_sorted)))),
  start.degree = 90,
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0)
)
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_BL_EukProk_atLeast3.pdf"))
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```

Chord Diagram: Prokaryote-Prokaryote EA
```{r}
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_prok_prok_EA$Source_sorted, interaction_table_prok_prok_EA$Target_sorted)))),
  start.degree = 90,  # 90 degrees moves 0° to the top
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0))
chordDiagram(
  interaction_table_prok_prok_EA %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_EA_ProkProk_all.pdf"))
chordDiagram(
  interaction_table_prok_prok_EA %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```
Only show edges with weight ≥ 3
```{r}
# Filter: Keep only interactions with weight >= 3
interaction_table_filtered <- interaction_table_prok_prok_EA %>%
  filter(Weight >= 3)

# Create Chord Diagram
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_filtered$Source_sorted, interaction_table_filtered$Target_sorted)))),
  start.degree = 90,
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0)
)
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_EA_ProkProk_atLeast3.pdf"))
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```

Chord Diagram: Eukaryote-Eukaryote EA
```{r}
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_euk_euk_EA$Source_sorted, interaction_table_euk_euk_EA$Target_sorted)))),
  start.degree = 90,  # 90 degrees moves 0° to the top
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0))
chordDiagram(
  interaction_table_euk_euk_EA %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_EA_EukEuk_all.pdf"))
chordDiagram(
  interaction_table_euk_euk_EA %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```

Only show edges with weight ≥ 3
```{r}
# Filter: Keep only interactions with weight >= 3
interaction_table_filtered <- interaction_table_euk_euk_EA %>%
  filter(Weight >= 3)

# Create Chord Diagram
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_filtered$Source_sorted, interaction_table_filtered$Target_sorted)))),
  start.degree = 90,
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0)
)
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_EA_EukEuk_atLeast3.pdf"))
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```

Chord Diagram: Prokaryote-Eukaryote EA
```{r}
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_cross_EA$Source_sorted, interaction_table_cross_EA$Target_sorted)))),
  start.degree = 90,  # 90 degrees moves 0° to the top
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0))
chordDiagram(
  interaction_table_cross_EA %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_EA_EukProk_all.pdf"))
chordDiagram(
  interaction_table_cross_EA %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```
Only show edges with weight ≥ 3
```{r}
# Filter: Keep only interactions with weight >= 3
interaction_table_filtered <- interaction_table_cross_EA %>%
  filter(Weight >= 3)

# Create Chord Diagram
circos.clear()
circos.par(
  gap.after = rep(5, length(unique(c(interaction_table_filtered$Source_sorted, interaction_table_filtered$Target_sorted)))),
  start.degree = 90,
  track.margin = c(0.05, 0.05),
  cell.padding = c(0, 0, 0, 0)
)
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)

# Save the plot
pdf(file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_EA_EukProk_atLeast3.pdf"))
chordDiagram(
  interaction_table_filtered %>% select(Source_sorted, Target_sorted, Weight),
  grid.col = grid_colors_all,
  transparency = 0.3,
  annotationTrack = "grid"
)
# Add labels
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  sector.name = get.cell.meta.data("sector.index")
  circos.text(
    x = mean(get.cell.meta.data("xlim")),
    y = 1.1,
    labels = abbreviation_map[sector.name],
    facing = "clockwise",
    niceFacing = TRUE,
    adj = c(0, 0.5),
    cex = 0.7
  )
}, bg.border = NA)
dev.off()

circos.clear()
```
Legend
```{r}
# Create a DataFrame for the legend
legend_df <- data.frame(
  Color = grid_colors_all[Groups_all],            # HEX color codes
  Abbreviation = abbreviation_map[Groups_all],    # 3-letter (or custom) abbreviation
  FullName = Groups_all,                          # Full taxonomic name
  stringsAsFactors = FALSE
)

# Save the table
write.csv(legend_df, file = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_Legend.csv"), row.names = FALSE)

# Display legend
kable(legend_df)
```

Color swatch legend
```{r}
# Create the legend table if not already
legend_df <- data.frame(
  Color = grid_colors_all[Groups_all],
  Abbreviation = abbreviation_map[Groups_all],
  FullName = Groups_all,
  stringsAsFactors = FALSE
)

# Create kable with color swatch
legend_df %>%
  mutate(Color = cell_spec(" ", background = Color)) %>%  # Create empty cell with background color
  kable(escape = FALSE, align = "c", format = "html") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed"))
```

Save as WORD
```{r}
# Legend table
legend_df <- data.frame(
  Color = grid_colors_all[Groups_all],
  Abbreviation = abbreviation_map[Groups_all],
  FullName = Groups_all,
  stringsAsFactors = FALSE
)

# Create flextable
ft <- flextable(legend_df)

# Set background color for the Color column
for (i in seq_len(nrow(legend_df))) {
  ft <- bg(ft, i = i, j = 1, bg = legend_df$Color[i])   # 1 is the first column = Color
}

# Set text color for Color column to white (optional, makes it clean)
ft <- color(ft, j = 1, color = "white")

# Autofit
ft <- autofit(ft)

# Show in Viewer
ft

# Save as Word
doc <- read_docx()
doc <- body_add_flextable(doc, ft)
print(doc, target = paste0(PATH, "/5_NetworkVisualization/ChordDiagram_StaticNW_Legend.docx"))
```
