---
title: "TemporalNetworks"
author: "Ina Maria Deutschmann"
date: "01/12/2025"
output: html_document
---

# Overview
1. Preprocessing
- remove and store taxonomic information in separate file
- filter for samples containing both kingdoms and both size fractions
- filter out rare ASVs (abundance sum > 100 reads and frequency > 15% occurrences)
- size fraction filtering as done in https://microbiomejournal.biomedcentral.com/articles/10.1186/s40168-023-01523-z (GitHub code: https://github.com/InaMariaDeutschmann/TemporalNetworkBBMO/blob/main/0_FromASVTableToTemporalNetwork.Rmd)
- clr-transformation
- merge tables to have one table per location (BL and EA)

2. Network inference
- Network construction with FlashWeave to infer single static network
- Removing environmental edges with EnDED
- Determine temporal network from single static network via subnetworks: an edge is present in a monthly subnetwork when both microorganisms of the predicted associations are present (abundance > 0)

## Set Up
Variables
```{r}
PATH <- "~/Documents/Ina/Pedro/"
CUTOFF_ABUNDANCE <- 100
CUTOFF_FREQUENCY <- 15

file_ASV16S <- "0_PedroData/ASVtab_16SV4V5_seqs_tax_v138_CLEANED_NoMock_NoChla_NoMitoch_min5570r.csv"
file_ASV18S <- "0_PedroData/ASVtab_18SV4_seqs_tax_pr2V4_14_NoMetazoaFungi_NoEukNA_NoStreptophyta_NoOpisthokontaNA_min9513.csv"
file_ENV <- "0_PedroData/EnvData_EAMO_BBMO_12112024.csv"
```

Required Packages
```{r, echo = FALSE}
library(dplyr)         # mutate_at
library(compositions)  # clr transform
library(VennDiagram)
library(RColorBrewer)
color_palette <- brewer.pal(4, "Set2")
```

## 1. Preprocessing
- remove and store taxonomic information in separate file
```{r}
# 16S
dt <- read.csv(paste(PATH, file_ASV16S, sep=""), header = TRUE)
colnames(dt)[1] <- "Sequence"
dt_temp <- dt[,c("ASV", "Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Sequence")]
colnames(dt_temp)[1] <- "ID"
write.table(dt_temp, paste(PATH,"1_Data/ASV_16S_seqs_tax.tsv",sep=""), col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
dt$Kingdom <- NULL
dt$Phylum <- NULL
dt$Class <- NULL
dt$Order <- NULL
dt$Family <- NULL
dt$Genus <- NULL
dt_16S <- dt[,sort(colnames(dt))]

# 18S
dt <- read.csv(paste(PATH, file_ASV18S, sep=""), header = TRUE)
colnames(dt)[1] <- "Sequence"
dt_temp <- dt[,c("ASV", "Kingdom", "Supergroup", "Division", "Class", "Order", "Family", "Genus", "Species", "reads", "Sequence")]
colnames(dt_temp)[1] <- "ID"
write.table(dt_temp, paste(PATH,"1_Data/ASV_18S_seqs_tax.tsv",sep=""), col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
dt$Kingdom <- NULL
dt$Supergroup <- NULL
dt$Division <- NULL
dt$Class <- NULL
dt$Order <- NULL
dt$Family <- NULL
dt$Genus <- NULL
dt$Species <- NULL
dt$reads <- NULL
dt$Sequence <- NULL
dt_18S <- dt[,sort(colnames(dt))]
```

-> number of samples
--> 16S: 122 samples
--> 18S: 130 samples

- separate samples based on sizefraction
```{r}
# sample IDs 16S
BL_samples <- colnames(dt_16S)[grep("BL",colnames(dt_16S))]
BL_16S_small_size_fraction <- sort(gsub(".16SV4V5.022.3.SRF","", BL_samples[grep("022.3", BL_samples)]))
BL_16S_big_size_fraction <- sort(gsub(".16SV4V5.3.200.SRF","", BL_samples[grep("3.200", BL_samples)]))

EA_samples <- colnames(dt_16S)[grep("EA",colnames(dt_16S))]
EA_16S_small_size_fraction <- sort(gsub(".16SV4V5.022.3.SRF","", EA_samples[grep("022.3", EA_samples)]))
EA_16S_big_size_fraction <- sort(gsub(".16SV4V5.3.200.SRF","", EA_samples[grep("3.200", EA_samples)]))

# sample IDs 18S
BL_samples <- colnames(dt_18S)[grep("BL",colnames(dt_18S))]
BL_18S_small_size_fraction <- sort(gsub(".18SV4.022.3.SRF","", BL_samples[grep("022.3", BL_samples)]))
BL_18S_big_size_fraction <- sort(gsub(".18SV4.3.200.SRF","", BL_samples[grep("3.200", BL_samples)]))

EA_samples <- colnames(dt_18S)[grep("EA",colnames(dt_18S))]
EA_18S_small_size_fraction <- sort(gsub(".18SV4.022.3.SRF","", EA_samples[grep("022.3", EA_samples)]))
EA_18S_big_size_fraction <- sort(gsub(".18SV4.3.200.SRF","", EA_samples[grep("3.200", EA_samples)]))
```

- Overview: how many samples (BL)
```{r}
# Create the data frame with all samples
dt_temp <- data.frame(ID = "BL_16S_022_3", samples = BL_16S_small_size_fraction)
dt_temp <- rbind(dt_temp,
                 data.frame(ID = "BL_16S_3_200", samples = BL_16S_big_size_fraction))
dt_temp <- rbind(dt_temp,
                 data.frame(ID = "BL_18S_022_3", samples = BL_18S_small_size_fraction))
dt_temp <- rbind(dt_temp,
                 data.frame(ID = "BL_18S_3_200", samples = BL_18S_big_size_fraction))

# Extract unique samples for each group
BL_16S_small <- unique(dt_temp$samples[dt_temp$ID == "BL_16S_022_3"])
BL_16S_big <- unique(dt_temp$samples[dt_temp$ID == "BL_16S_3_200"])
BL_18S_small <- unique(dt_temp$samples[dt_temp$ID == "BL_18S_022_3"])
BL_18S_big <- unique(dt_temp$samples[dt_temp$ID == "BL_18S_3_200"])

# Create a list of these sample sets
venn_data <- list(
  "16S Small" = BL_16S_small,
  "16S Big" = BL_16S_big,
  "18S Small" = BL_18S_small,
  "18S Big" = BL_18S_big
)

# Generate the Venn diagram
venn.plot <- venn.diagram(
  x = venn_data,
  category.names = names(venn_data),
  filename = NULL,  # Keep this NULL to display it directly
  output = TRUE,
  
  # Styling options for the diagram
  lwd = 2,
  fill = color_palette,
  alpha = 0.5,
  cex = 1.5,
  fontfamily = "sans",
  cat.fontfamily = "sans",
  cat.cex = 1.5,
  cat.pos = 0
)

# Save the Venn diagram as a PDF file
pdf("1_Data/venn_diagram_NumSamples_BL.pdf", width = 8, height = 8)  # Opens a PDF device
grid.newpage()  # Starts a new page in the PDF
grid.draw(venn.plot)  # Draws the Venn diagram on the PDF page
dev.off()  # Closes the PDF device and saves the file
```

- Venn Diagram BL: How many samples contain the kingdoms and size fractions
```{r, echo=FALSE, fig.width=3, fig.height=3, fig.align='center'}
# Plot the Venn diagram
grid.newpage()
grid.draw(venn.plot)
```

- Table BL: Which samples contain which kingdoms and size fractions
```{r}
t(table(dt_temp))
```
             ID
samples       BL_16S_022_3 BL_16S_3_200 BL_18S_022_3 BL_18S_3_200
  BL.20130206            1            0            0            0
  BL.20130417            1            1            1            1
  BL.20130507            1            1            1            1
  BL.20130604            1            1            1            1
  BL.20130709            1            1            1            1
  BL.20130801            1            1            1            1
  BL.20130917            1            1            1            1
  BL.20131015            1            1            1            1
  BL.20131105            1            1            1            0
  BL.20131204            1            1            1            0
  BL.20140114            1            1            1            1
  BL.20140211            1            1            1            1
  BL.20140310            1            1            1            1
  BL.20140407            0            1            0            1
  BL.20140505            1            1            0            1
  BL.20140602            1            1            0            1
  BL.20140621            1            0            0            1
  BL.20140707            1            0            0            1
  BL.20140804            1            0            0            1
  BL.20140916            1            0            0            1
  BL.20141007            1            1            1            1
  BL.20141111            1            1            1            1
  BL.20141216            1            0            1            1
  BL.20150113            1            1            1            1
  BL.20150210            1            1            1            1
  BL.20150310            1            0            1            1
  BL.20150415            1            1            0            1
  BL.20150512            1            0            1            1
  BL.20150609            0            0            1            1
  BL.20150622            0            0            1            1
  BL.20150707            1            0            1            1
  BL.20150804            1            0            1            1
  BL.20150916            1            1            1            1
  BL.20151006            1            1            1            1
  BL.20151110            1            1            1            1
  BL.20151215            1            1            1            1
  BL.20160119            1            1            1            1
  BL.20160210            1            1            1            1
  BL.20160308            1            1            1            1
  BL.20160405            1            1            1            1
  BL.20160504            1            1            1            1
  BL.20160607            1            0            1            1
  BL.20160802            1            1            0            0

- Overview: how many samples (EA)
```{r}
# Create the data frame with all samples
dt_temp <- data.frame(ID = "EA_16S_022_3", samples = EA_16S_small_size_fraction)
dt_temp <- rbind(dt_temp,
                 data.frame(ID = "EA_16S_3_200", samples = EA_16S_big_size_fraction))
dt_temp <- rbind(dt_temp,
                 data.frame(ID = "EA_18S_022_3", samples = EA_18S_small_size_fraction))
dt_temp <- rbind(dt_temp,
                 data.frame(ID = "EA_18S_3_200", samples = EA_18S_big_size_fraction))

# Extract unique samples for each group
EA_16S_small <- unique(dt_temp$samples[dt_temp$ID == "EA_16S_022_3"])
EA_16S_big <- unique(dt_temp$samples[dt_temp$ID == "EA_16S_3_200"])
EA_18S_small <- unique(dt_temp$samples[dt_temp$ID == "EA_18S_022_3"])
EA_18S_big <- unique(dt_temp$samples[dt_temp$ID == "EA_18S_3_200"])

# Create a list of these sample sets
venn_data <- list(
  "16S Small" = EA_16S_small,
  "16S Big" = EA_16S_big,
  "18S Small" = EA_18S_small,
  "18S Big" = EA_18S_big
)

# Generate the Venn diagram
venn.plot <- venn.diagram(
  x = venn_data,
  category.names = names(venn_data),
  filename = NULL,  # Keep this NULL to display it directly
  output = TRUE,
  
  # Styling options for the diagram
  lwd = 2,
  fill = color_palette,
  alpha = 0.5,
  cex = 1.5,
  fontfamily = "sans",
  cat.fontfamily = "sans",
  cat.cex = 1.5,
  cat.pos = 0
)

# Save the Venn diagram as a PDF file
pdf("1_Data/venn_diagram_NumSamples_EA.pdf", width = 8, height = 8)  # Opens a PDF device
grid.newpage()  # Starts a new page in the PDF
grid.draw(venn.plot)  # Draws the Venn diagram on the PDF page
dev.off()  # Closes the PDF device and saves the file
```

- Venn Diagram EA: How many samples contain the kingdoms and size fractions
```{r, echo=FALSE, fig.width=3, fig.height=3, fig.align='center'}
# Plot the Venn diagram
grid.newpage()
grid.draw(venn.plot)
```

- Table EA: Which samples contain which kingdoms and size fractions
```{r}
t(table(dt_temp))
```
             ID
samples       EA_16S_022_3 EA_16S_3_200 EA_18S_022_3 EA_18S_3_200
  EA.20130423            1            1            1            1
  EA.20130604            1            1            1            1
  EA.20130625            1            1            1            1
  EA.20130827            1            0            1            0
  EA.20131127            1            1            1            1
  EA.20131230            1            1            1            1
  EA.20140211            1            0            1            1
  EA.20140326            1            0            0            1
  EA.20140415            1            1            1            1
  EA.20140514            1            1            1            1
  EA.20140707            1            0            0            1
  EA.20140729            1            0            1            0
  EA.20140826            1            0            1            1
  EA.20140930            1            0            1            1
  EA.20141029            1            1            1            1
  EA.20141125            1            0            1            1
  EA.20150115            0            0            1            1
  EA.20150227            0            0            1            1
  EA.20150228            1            1            0            0
  EA.20150328            1            1            1            1
  EA.20150425            1            0            1            1
  EA.20150516            1            1            1            1
  EA.20150627            1            1            1            1
  EA.20150801            1            1            1            1
  EA.20150829            1            1            1            1
  EA.20151006            1            1            1            1
  EA.20151208            1            1            1            1
  EA.20160225            1            1            1            1
  EA.20160322            1            1            1            1
  EA.20160502            1            1            1            1
  EA.20160611            1            1            1            1
  EA.20160822            1            1            1            1

-> 23 samples BL and 22 samples EA have 16S and 18S in both size fractions

Save
```{r}
write.table(dt_16S, paste(PATH,"1_Data/ASV_16S_all_abund.tsv",sep=""), col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
write.table(dt_18S, paste(PATH,"1_Data/ASV_18S_all_abund.tsv",sep=""), col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
```

- Overview number of samples per location, kingdom, and size fraction
```{r}
# Read in tables
dt_16S <- read.table(paste(PATH,"1_Data/ASV_16S_all_abund.tsv",sep=""), header = TRUE, sep = "\t", stringsAsFactors = FALSE)
dt_18S <- read.table(paste(PATH,"1_Data/ASV_18S_all_abund.tsv",sep=""), header = TRUE, sep = "\t", stringsAsFactors = FALSE)

# Number of samples after sample filtering
print("BL")
print(paste("16S_022_3:",length(BL_16S_small_size_fraction)))
print(paste("16S_3_200:",length(BL_16S_big_size_fraction)))
print(paste("16S_022_3 + 16S_3_200:",length(intersect(BL_16S_small_size_fraction, BL_16S_big_size_fraction))))
print(paste("18S_022_3:",length(BL_18S_small_size_fraction)))
print(paste("18S_3_200:",length(BL_18S_big_size_fraction)))
print(paste("18S_022_3 + 18S_3_200:",length(intersect(BL_18S_small_size_fraction, BL_18S_big_size_fraction))))
print(paste("16S_022_3 + 18S_022_3:",length(intersect(BL_16S_small_size_fraction, BL_18S_small_size_fraction))))
print(paste("16S_3_200 + 18S_3_200:",length(intersect(BL_16S_big_size_fraction, BL_18S_big_size_fraction))))
print(paste("all 4:",length(intersect(intersect(BL_16S_small_size_fraction, BL_16S_big_size_fraction), intersect(BL_18S_small_size_fraction, BL_18S_big_size_fraction)))))

print("EA")
print(paste("16S_022_3:",length(EA_16S_small_size_fraction)))
print(paste("16S_3_200:",length(EA_16S_big_size_fraction)))
print(paste("16S_022_3 + 16S_3_200:",length(intersect(EA_16S_small_size_fraction, EA_16S_big_size_fraction))))
print(paste("18S_022_3:",length(EA_18S_small_size_fraction)))
print(paste("18S_3_200:",length(EA_18S_big_size_fraction)))
print(paste("18S_022_3 + 18S_3_200:",length(intersect(EA_18S_small_size_fraction, EA_18S_big_size_fraction))))
print(paste("16S_022_3 + 18S_022_3:",length(intersect(EA_16S_small_size_fraction, EA_18S_small_size_fraction))))
print(paste("16S_3_200 + 18S_3_200:",length(intersect(EA_16S_big_size_fraction, EA_18S_big_size_fraction))))
print(paste("all 4:",length(intersect(intersect(EA_16S_small_size_fraction, EA_16S_big_size_fraction), intersect(EA_18S_small_size_fraction, EA_18S_big_size_fraction)))))
```
[1] "BL"
[1] "16S_022_3: 40"
[1] "16S_3_200: 30"
[1] "16S_022_3 + 16S_3_200: 29"
[1] "18S_022_3: 33"
[1] "18S_3_200: 39"
[1] "18S_022_3 + 18S_3_200: 31"
[1] "16S_022_3 + 18S_022_3: 31"
[1] "16S_3_200 + 18S_3_200: 27"
[1] "all 4: 23"
[1] "EA"
[1] "16S_022_3: 30"
[1] "16S_3_200: 22"
[1] "16S_022_3 + 16S_3_200: 22"
[1] "18S_022_3: 29"
[1] "18S_3_200: 29"
[1] "18S_022_3 + 18S_3_200: 27"
[1] "16S_022_3 + 18S_022_3: 28"
[1] "16S_3_200 + 18S_3_200: 22"
[1] "all 4: 22"

- filter out rare ASVs: abundance sum above CUTOFF_ABUNDANCE and prevalence above CUTOFF_FREQUENCY
```{r}
# Overview
# 16S
print("16S")
for(i in c("BL", "EA")){
  for(j in c("022.3", "3.200")){
    dt <- dt_16S[,grep(i,colnames(dt_16S))]
    dt <- dt[,grep(j,colnames(dt))]
    print(paste(i,j))
    print(paste(sum(rowSums(dt)>CUTOFF_ABUNDANCE), " ASVs with abundance sum > ", CUTOFF_ABUNDANCE, sep=""))
    print(paste(sum((rowSums(1*(dt>0))/(dim(dt)[2])*100)>CUTOFF_FREQUENCY), " ASVs with prevalence > ", CUTOFF_FREQUENCY, "% (present in >",(dim(dt)[2])/100*CUTOFF_FREQUENCY," samples)", sep=""))
    dt_temp <- dt[which((rowSums(dt)>CUTOFF_ABUNDANCE) & ((rowSums(1*(dt>0))/(dim(dt)[2])*100)>CUTOFF_FREQUENCY)),]
    print(paste(dim(dt_temp)[1], " ASVs with abundance sum > ", CUTOFF_ABUNDANCE, " and prevalence > ", CUTOFF_FREQUENCY, "% (present in >",(dim(dt)[2])/100*CUTOFF_FREQUENCY," samples)", sep=""))
  }
}
# 18S
print("18S")
for(i in c("BL", "EA")){
  for(j in c("022.3", "3.200")){
    dt <- dt_18S[,grep(i,colnames(dt_16S))]
    dt <- dt[,grep(j,colnames(dt))]
    print(paste(i,j))
    print(paste(sum(rowSums(dt)>CUTOFF_ABUNDANCE), " ASVs with abundance sum > ", CUTOFF_ABUNDANCE, sep=""))
    print(paste(sum((rowSums(1*(dt>0))/(dim(dt)[2])*100)>CUTOFF_FREQUENCY), " ASVs with prevalence > ", CUTOFF_FREQUENCY, "% (present in >",(dim(dt)[2])/100*CUTOFF_FREQUENCY," samples)", sep=""))
    dt_temp <- dt[which((rowSums(dt)>CUTOFF_ABUNDANCE) & ((rowSums(1*(dt>0))/(dim(dt)[2])*100)>CUTOFF_FREQUENCY)),]
    print(paste(dim(dt_temp)[1], " ASVs with abundance sum > ", CUTOFF_ABUNDANCE, " and prevalence > ", CUTOFF_FREQUENCY, "% (present in >",(dim(dt)[2])/100*CUTOFF_FREQUENCY," samples)", sep=""))
  }
}

# Create tables
BL_intersection <- intersect(intersect(BL_16S_small_size_fraction, BL_16S_big_size_fraction),
                             intersect(BL_18S_small_size_fraction, BL_18S_big_size_fraction))
EA_intersection <- intersect(intersect(EA_16S_small_size_fraction, EA_16S_big_size_fraction),
                             intersect(EA_18S_small_size_fraction, EA_18S_big_size_fraction))

for(i in c("BL", "EA"))
{
  if(i=="BL"){
    col_names <- BL_intersection
  }else{
    col_names <- EA_intersection
  }
  for(j in c("022.3", "3.200"))
  {
    # 16S
    dt <- dt_16S[,grep(i,colnames(dt_16S))]
    dt <- dt[,grep(j,colnames(dt))]
    rownames(dt) <- dt_16S$ASV
    colnames(dt) <- gsub(".18SV4.", "", gsub(".16SV4V5.", "", gsub("022.3.SRF","", gsub("3.200.SRF","", colnames(dt)))))
    dt <- dt[,col_names]
    print(paste(i, j, "16S before filtering:", dim(dt)[1], dim(dt)[2]))
    dt <- dt[which((rowSums(dt)>CUTOFF_ABUNDANCE) & ((rowSums(1*(dt>0))/(dim(dt)[2])*100)>CUTOFF_FREQUENCY)),]
    print(paste(i, j, "16S after filtering:", dim(dt)[1], dim(dt)[2]))
    dt$ASV <- paste(rownames(dt),"16S",j,sep="_")
    dt <- dt[,sort(colnames(dt))]
    rownames(dt) <- dt$ASV
    write.table(dt, paste(PATH,"1_Data/",i,"_",j,"_16S_abundances_rareASV_filtering.tsv", sep=""), col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
    print(paste("16S + ", i, " + ", j, " (intersection of samples): ", dim(dt)[1], " ASVs with abundance sum > ", CUTOFF_ABUNDANCE, " and prevalence > ", CUTOFF_FREQUENCY, "% (present in >",(dim(dt)[2])/100*CUTOFF_FREQUENCY," samples)", sep=""))
    
    # 18S
    dt <- dt_18S[,grep(i,colnames(dt_18S))]
    dt <- dt[,grep(j,colnames(dt))]
    rownames(dt) <- dt_18S$ASV
    colnames(dt) <- gsub(".18SV4.", "", gsub(".16SV4V5.", "", gsub("022.3.SRF","", gsub("3.200.SRF","", colnames(dt)))))
    dt <- dt[,col_names]
    print(paste(i, j, "18S before filtering:", dim(dt)[1], dim(dt)[2]))
    dt <- dt[which((rowSums(dt)>CUTOFF_ABUNDANCE) & ((rowSums(1*(dt>0))/(dim(dt)[2])*100)>CUTOFF_FREQUENCY)),]
    print(paste(i, j, "18S after filtering:", dim(dt)[1], dim(dt)[2]))
    dt$ASV <- paste(rownames(dt),"18S",j,sep="_")
    dt <- dt[,sort(colnames(dt))]
    rownames(dt) <- dt$ASV
    write.table(dt, paste(PATH,"1_Data/",i,"_",j,"_18S_abundances_rareASV_filtering.tsv", sep=""), col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
    print(paste("18S + ", i, " + ", j, " (intersection of samples): ", dim(dt)[1], " ASVs with abundance sum > ", CUTOFF_ABUNDANCE, " and prevalence > ", CUTOFF_FREQUENCY, "% (present in >",(dim(dt)[2])/100*CUTOFF_FREQUENCY," samples)", sep=""))

  }
}
```
-> with abundance cut_off 100, and frequency 15%
[1] "16S"
[1] "BL 022.3"
[1] "1137 ASVs with abundance sum > 100"
[1] "981 ASVs with prevalence > 15% (present in >6 samples)"
[1] "818 ASVs with abundance sum > 100 and prevalence > 15% (present in >6 samples)"

[1] "BL 3.200"
[1] "578 ASVs with abundance sum > 100"
[1] "617 ASVs with prevalence > 15% (present in >4.5 samples)"
[1] "442 ASVs with abundance sum > 100 and prevalence > 15% (present in >4.5 samples)"

[1] "EA 022.3"
[1] "831 ASVs with abundance sum > 100"
[1] "924 ASVs with prevalence > 15% (present in >4.5 samples)"
[1] "655 ASVs with abundance sum > 100 and prevalence > 15% (present in >4.5 samples)"

[1] "EA 3.200"
[1] "308 ASVs with abundance sum > 100"
[1] "507 ASVs with prevalence > 15% (present in >3.3 samples)"
[1] "257 ASVs with abundance sum > 100 and prevalence > 15% (present in >3.3 samples)"

[1] "18S"
[1] "BL 022.3"
[1] "2547 ASVs with abundance sum > 100"
[1] "662 ASVs with prevalence > 15% (present in >4.8 samples)"
[1] "628 ASVs with abundance sum > 100 and prevalence > 15% (present in >4.8 samples)"

[1] "BL 3.200"
[1] "2395 ASVs with abundance sum > 100"
[1] "619 ASVs with prevalence > 15% (present in >5.7 samples)"
[1] "598 ASVs with abundance sum > 100 and prevalence > 15% (present in >5.7 samples)"

[1] "EA 022.3"
[1] "2358 ASVs with abundance sum > 100"
[1] "1657 ASVs with prevalence > 15% (present in >3.9 samples)"
[1] "1447 ASVs with abundance sum > 100 and prevalence > 15% (present in >3.9 samples)"

[1] "EA 3.200"
[1] "2024 ASVs with abundance sum > 100"
[1] "1220 ASVs with prevalence > 15% (present in >3.9 samples)"
[1] "1083 ASVs with abundance sum > 100 and prevalence > 15% (present in >3.9 samples)"

[1] "BL 022.3 16S before filtering: 20335 23"
[1] "BL 022.3 16S after filtering: 797 23"
[1] "16S + BL + 022.3 (intersection of samples): 797 ASVs with abundance sum > 100 and prevalence > 15% (present in >3.6 samples)"

[1] "BL 022.3 18S before filtering: 20699 23"
[1] "BL 022.3 18S after filtering: 577 23"
[1] "18S + BL + 022.3 (intersection of samples): 577 ASVs with abundance sum > 100 and prevalence > 15% (present in >3.6 samples)"

[1] "BL 3.200 16S before filtering: 20335 23"
[1] "BL 3.200 16S after filtering: 414 23"
[1] "16S + BL + 3.200 (intersection of samples): 414 ASVs with abundance sum > 100 and prevalence > 15% (present in >3.6 samples)"

[1] "BL 3.200 18S before filtering: 20699 23"
[1] "BL 3.200 18S after filtering: 690 23"
[1] "18S + BL + 3.200 (intersection of samples): 690 ASVs with abundance sum > 100 and prevalence > 15% (present in >3.6 samples)"

[1] "EA 022.3 16S before filtering: 20335 22"
[1] "EA 022.3 16S after filtering: 578 22"
[1] "16S + EA + 022.3 (intersection of samples): 578 ASVs with abundance sum > 100 and prevalence > 15% (present in >3.45 samples)"

[1] "EA 022.3 18S before filtering: 20699 22"
[1] "EA 022.3 18S after filtering: 1296 22"
[1] "18S + EA + 022.3 (intersection of samples): 1296 ASVs with abundance sum > 100 and prevalence > 15% (present in >3.45 samples)"

[1] "EA 3.200 16S before filtering: 20335 22"
[1] "EA 3.200 16S after filtering: 257 22"
[1] "16S + EA + 3.200 (intersection of samples): 257 ASVs with abundance sum > 100 and prevalence > 15% (present in >3.45 samples)"

[1] "EA 3.200 18S before filtering: 20699 22"
[1] "EA 3.200 18S after filtering: 1114 22"
[1] "18S + EA + 3.200 (intersection of samples): 1114 ASVs with abundance sum > 100 and prevalence > 15% (present in >3.45 samples)"

- Size fraction filtering as done in https://microbiomejournal.biomedcentral.com/articles/10.1186/s40168-023-01523-z (GitHub code: https://github.com/InaMariaDeutschmann/TemporalNetworkBBMO/blob/main/0_FromASVTableToTemporalNetwork.Rmd)
```{r}
for(i in c("BL", "EA"))
{
  for(k in c("16S", "18S"))
  {
    # Read in both size fraction tables
    big_df <- read.table(paste(PATH, "1_Data/", i, "_3.200_", k, "_abundances_rareASV_filtering.tsv", sep=""), header=TRUE, sep="\t")
    small_df <- read.table(paste(PATH, "1_Data/", i, "_022.3_", k, "_abundances_rareASV_filtering.tsv", sep=""), header=TRUE, sep="\t")
    
    # Perform size fraction filtering
    big_df$ASV <- gsub("_3.200", "", big_df$ASV)
    small_df$ASV <- gsub("_022.3", "", small_df$ASV)
    big_df$abundancesum_big <- rowSums(big_df[, -1])
    small_df$abundancesum_small <- rowSums(small_df[, -1])
    
    sizefractionfilter <- merge(small_df[, c("ASV", "abundancesum_small")], big_df[, c("ASV", "abundancesum_big")], by="ASV")
    sizefractionfilter$big_dividedby_small <- sizefractionfilter$abundancesum_big / sizefractionfilter$abundancesum_small
    
    # Size filtering based on ratio thresholds
    big_df_filtered <- big_df
    small_df_filtered <- small_df
    
    big_df_filtered[big_df_filtered$ASV %in% sizefractionfilter$ASV[sizefractionfilter$big_dividedby_small < 0.5], -1] <- 0
    small_df_filtered[small_df_filtered$ASV %in% sizefractionfilter$ASV[sizefractionfilter$big_dividedby_small > 2], -1] <- 0
    
    big_df_filtered <- big_df_filtered[rowSums(big_df_filtered[, -1]) > 0, ]
    small_df_filtered <- small_df_filtered[rowSums(small_df_filtered[, -1]) > 0, ]
    
    # Overview of ASVs removed
    big_removed <- nrow(sizefractionfilter[sizefractionfilter$big_dividedby_small < 0.5, ])
    small_removed <- nrow(sizefractionfilter[sizefractionfilter$big_dividedby_small > 2, ])
    
    cat("Overview of ASV removal in ", i, " + ", k, ":\n")
    cat("ASVs removed from big size fraction (<0.5):", big_removed, "\n")
    cat("ASVs remaining in big size fraction:", dim(big_df_filtered)[1], "\n")
    cat("ASVs removed from small size fraction (>2):", small_removed, "\n")
    cat("ASVs remaining in small size fraction:", dim(small_df_filtered)[1], "\n")
    
    # Save filtered tables
    write.table(
      big_df_filtered, 
      paste(PATH, "1_Data/", i, "_3.200_", k, "_abundances_rareASVandSizeFraction_filtering.tsv", sep=""), 
      col.names=TRUE, row.names=FALSE, sep="\t", quote=FALSE
    )
    
    write.table(
      small_df_filtered, 
      paste(PATH, "1_Data/", i, "_022.3_", k, "_abundances_rareASVandSizeFraction_filtering.tsv", sep=""), 
      col.names=TRUE, row.names=FALSE, sep="\t", quote=FALSE
    )
  }
}
```

Overview of ASV removal in  BL  +  16S :
ASVs removed from big size fraction (<0.5): 239 
ASVs remaining in big size fraction: 175 
ASVs removed from small size fraction (>2): 19 
ASVs remaining in small size fraction: 778 

Overview of ASV removal in  BL  +  18S :
ASVs removed from big size fraction (<0.5): 146 
ASVs remaining in big size fraction: 544 
ASVs removed from small size fraction (>2): 43 
ASVs remaining in small size fraction: 534 

Overview of ASV removal in  EA  +  16S :
ASVs removed from big size fraction (<0.5): 117 
ASVs remaining in big size fraction: 140 
ASVs removed from small size fraction (>2): 14 
ASVs remaining in small size fraction: 564 

Overview of ASV removal in  EA  +  18S :
ASVs removed from big size fraction (<0.5): 339 
ASVs remaining in big size fraction: 775 
ASVs removed from small size fraction (>2): 208 
ASVs remaining in small size fraction: 1088 

- clr transformation + save file (each table)
```{r}
# Function to compute Central log ratio transformation and return result as a vector
my_clr <- function(x) {as.vector(clr(x))}
pseudo_count <- 1

for(i in c("BL", "EA"))
{
  for(j in c("022.3", "3.200"))
  {
    for(k in c("16S", "18S"))
    {
      ## Read in table
      dt <- read.table(paste(PATH, "1_Data/", i, "_", j, "_", k, "_abundances_rareASVandSizeFraction_filtering.tsv", sep=""), header=TRUE, sep="\t")
      dt$abundancesum_big <- NULL
      dt$abundancesum_small <- NULL

      ## Add pseudo count of 1 and Central log ratio transformation
      ASV_names <- dt$ASV
      dt$ASV <- NULL
      dt_temp_clr <- dplyr::mutate_at((dt + pseudo_count), colnames(dt), my_clr)
      dt_temp_clr$ASV <- ASV_names
      dt_temp_clr <- dt_temp_clr[,sort(colnames(dt_temp_clr))]
      write.table(dt_temp_clr, paste(PATH,"1_Data/",i,"_", j, "_", k, "_clr.tsv", sep=""), col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
    }
  }
}
```

- merge tables that are already filtered and clr transformed + save file
```{r}
for(i in c("BL", "EA"))
{
  if(i=="BL"){
    col_names <- c("ASV",BL_intersection)
  }else{
    col_names <- c("ASV",EA_intersection)
  }
  dt_022_3_16S <- read.table(paste(PATH,"1_Data/",i,"_022.3_16S_clr.tsv", sep=""), header = TRUE, sep = "\t")
  dt_022_3_16S$ASV <- gsub("_16S", "_16S_022_3", dt_022_3_16S$ASV)
  global_min <- min(dt_022_3_16S[,-1])
  
  dt_022_3_18S <- read.table(paste(PATH,"1_Data/",i,"_022.3_18S_clr.tsv", sep=""), header = TRUE, sep = "\t")
  dt_022_3_18S$ASV <- gsub("_18S", "_18S_022_3", dt_022_3_18S$ASV)
  if(global_min > min(dt_022_3_18S[,-1])){
    global_min <- min(dt_022_3_18S[,-1])
  }
  
  dt_3_200_16S <- read.table(paste(PATH,"1_Data/",i,"_3.200_16S_clr.tsv", sep=""), header = TRUE, sep = "\t")
  dt_3_200_16S$ASV <- gsub("_16S", "_16S_3_200", dt_3_200_16S$ASV)
  if(global_min > min(dt_3_200_16S[,-1])){
    global_min <- min(dt_3_200_16S[,-1])
  }
  
  dt_3_200_18S <- read.table(paste(PATH,"1_Data/",i,"_3.200_18S_clr.tsv", sep=""), header = TRUE, sep = "\t")
  dt_3_200_18S$ASV <- gsub("_18S", "_18S_3_200", dt_3_200_18S$ASV)
  if(global_min > min(dt_3_200_18S[,-1])){
    global_min <- min(dt_3_200_18S[,-1])
  }
  
  # Every local min (showing where zero was) will be set to the global min
  for(j in c(2:dim(dt_022_3_16S)[2]))
  {
    col_min = min(dt_022_3_16S[,j])
    dt_022_3_16S[which(dt_022_3_16S[,j]==col_min),j] <- global_min
  }
  for(j in c(2:dim(dt_022_3_18S)[2]))
  {
    col_min = min(dt_022_3_18S[,j])
    dt_022_3_18S[which(dt_022_3_18S[,j]==col_min),j] <- global_min
  }
  for(j in c(2:dim(dt_3_200_16S)[2]))
  {
    col_min = min(dt_3_200_16S[,j])
    dt_3_200_16S[which(dt_3_200_16S[,j]==col_min),j] <- global_min
  }
  for(j in c(2:dim(dt_3_200_18S)[2]))
  {
    col_min = min(dt_3_200_18S[,j])
    dt_3_200_18S[which(dt_3_200_18S[,j]==col_min),j] <- global_min
  }
  
  # Write tables
  ## One network per dataset (BL and EA), i.e., we merge kingdom (166S and 18S) and size fraction (022_3 and 3_200)
  write.table(rbind(dt_022_3_16S[,col_names], dt_022_3_18S[,col_names], dt_3_200_16S[,col_names], dt_3_200_18S[,col_names]),
              paste(PATH,"1_Data/",i,"_table_for_network_construction.tsv",sep=""), col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
}
```

## 2. Network inference

### 2.1. Network construction with FlashWeave
- Programming Language: julia version 1.10.5 (2024-08-27)
- FlashWeave to infer single static network (Version: FlashWeave v0.19.2)
- normalize=false because we already applied clr-transformation in order to merge data tables
- n_obs_min=3 because we require at least 3 observations
- make sure you are in folder "2_Networks/"
```{Julia}
import Pkg; Pkg.add("FlashWeave")
using FlashWeave
pkgs = Pkg.installed();
pkgs["FlashWeave"]

# BL
file_name="../1_Data/BL_table_for_network_construction.tsv"
netw_results = learn_network(file_name, sensitive=true, heterogeneous=true, normalize=false, transposed=true, n_obs_min=3)
save_network(replace(replace(file_name, "1_Data" => "2_Networks"), "table_for_network_construction.tsv" => "FlashWeave_network.edgelist"), netw_results)

# EA
file_name="../1_Data/EA_table_for_network_construction.tsv"
netw_results = learn_network(file_name, sensitive=true, heterogeneous=true, normalize=false, transposed=true, n_obs_min=3)
save_network(replace(replace(file_name, "1_Data" => "2_Networks"), "table_for_network_construction.tsv" => "FlashWeave_network.edgelist"), netw_results)
```

Output network construction BL
```
> netw_results = learn_network(file_name, sensitive=true, heterogeneous=true, normalize=false, transposed=true, n_obs_min=3)

### Loading data ###

┌ Warning: Skipping normalization, only experts should choose this option
└ @ FlashWeave ~/.julia/packages/FlashWeave/j91Ng/src/learning.jl:508
### Learning interactions ###

Inferring network with FlashWeaveHE - sensitive (conditional)

	Run information:
	sensitive - true
	heterogeneous - true
	max_k - 3
	alpha - 0.01
	sparse - true
	workers - 1
	OTUs - 2031
	MVs - 0

Computing univariate associations

Univariate degree stats:
Summary Stats:
Length:         2031
Missing Count:  0
Mean:           32.129985
Minimum:        0.000000
1st Quartile:   4.000000
Median:         18.000000
3rd Quartile:   42.000000
Maximum:        202.000000



Starting conditioning search

Preparing workers..

Done. Starting inference..
Starting convergence checks at 2347 edges.
Latest convergence step change: 0.93904
Latest convergence step change: 0.32367

Postprocessing
Complete

Finished inference. Total time taken: 6.483s

Mode:
FlashWeaveHE - sensitive (conditional)

Network:
2660 interactions between 2031 variables (2031 OTUs and 0 MVs)

Unfinished variables:
none

Rejections:
not tracked
```

Output network construction EA
```
> netw_results = learn_network(file_name, sensitive=true, heterogeneous=true, normalize=false, transposed=true, n_obs_min=3)

### Loading data ###

┌ Warning: Skipping normalization, only experts should choose this option
└ @ FlashWeave ~/.julia/packages/FlashWeave/j91Ng/src/learning.jl:508
### Learning interactions ###

Inferring network with FlashWeaveHE - sensitive (conditional)

	Run information:
	sensitive - true
	heterogeneous - true
	max_k - 3
	alpha - 0.01
	sparse - true
	workers - 1
	OTUs - 2567
	MVs - 0

Computing univariate associations

Univariate degree stats:
Summary Stats:
Length:         2567
Missing Count:  0
Mean:           12.508765
Minimum:        0.000000
1st Quartile:   1.000000
Median:         3.000000
3rd Quartile:   9.000000
Maximum:        187.000000



Starting conditioning search

Preparing workers..

Done. Starting inference..
Starting convergence checks at 2026 edges.
Latest convergence step change: 0.35537
Latest convergence step change: 0.26245
Latest convergence step change: 0.2684
Latest convergence step change: 0.21204
Latest convergence step change: 0.22998
Latest convergence step change: 0.20091
Latest convergence step change: 0.19861

Postprocessing
Complete

Finished inference. Total time taken: 1.379s

Mode:
FlashWeaveHE - sensitive (conditional)

Network:
2435 interactions between 2567 variables (2567 OTUs and 0 MVs)

Unfinished variables:
none

Rejections:
not tracked
```

=====> To keep in mind:
---> different number of samples
----> difficult/biased to compare networks because number of samples was not equal and it can bias network construction
-> However, 20 vs. 23 is similar!

- Number nodes with at least one edge (not considering isolated nodes)
```{r}
for(i in c("BL", "EA"))
{
  # Read in network
  nw <- read.table(paste(PATH,"2_Networks/", i, "_FlashWeave_network.edgelist", sep=""), header = FALSE, sep = "\t")
  
  # Rename columns in nw
  colnames(nw) <- c("Source", "Target", "association_score")
  
  # Combine node1 and node2, then find unique nodes
  all_nodes <- unique(c(nw$Source, nw$Target))
  
  # Print or save the results
  print(i)
  print(paste("#All nodes: ", length(all_nodes), sep=""))
  print(paste("#Nodes with 16S_022_3: ", length(all_nodes[grepl("16S_022_3", all_nodes)]), sep=""))
  print(paste("#Nodes with 16S_3_200: ", length(all_nodes[grepl("16S_3_200", all_nodes)]), sep=""))
  print(paste("#Nodes with 18S_022_3: ", length(all_nodes[grepl("18S_022_3", all_nodes)]), sep=""))
  print(paste("#Nodes with 18S_3_200: ", length(all_nodes[grepl("18S_3_200", all_nodes)]), sep=""))
  
  # Write table
  write.table(all_nodes, paste(PATH,"2_Networks/", i, "_all_nodes.txt", sep=""), col.names = FALSE, row.names = FALSE, quote = FALSE, sep = "\t")
  write.table(nw, paste(PATH,"2_Networks/", i, "_FlashWeave_network.tsv", sep=""), col.names = TRUE, row.names = FALSE, quote = FALSE, sep = "\t") 
}
```
[1] "BL"
[1] "#All nodes: 1914"
[1] "#Nodes with 16S_022_3: 756"
[1] "#Nodes with 16S_3_200: 159"
[1] "#Nodes with 18S_022_3: 468"
[1] "#Nodes with 18S_3_200: 531"

[1] "EA"
[1] "#All nodes: 2099"
[1] "#Nodes with 16S_022_3: 413"
[1] "#Nodes with 16S_3_200: 87"
[1] "#Nodes with 18S_022_3: 920"
[1] "#Nodes with 18S_3_200: 679"

### 2.2. Removing environmental edges with EnDED
- download EnDED (https://github.com/InaMariaDeutschmann/EnDED)
- on Mac in terminal
```{bash}
/bin/bash -c "$(curl -fsSL https://raw.githubusercontent.com/Homebrew/install/HEAD/install.sh)"

brew install gcc cmake
brew install boost
```

- added in EnDED makefile: "-I/opt/homebrew/opt/boost/include" in CXXFLAGS
```
CXXFLAGS := -std=c++11 -O3 -I/opt/homebrew/opt/boost/include
```

- prepare ASV abundance tables
```{r}
for(i in c("BL", "EA"))
{
  if(i=="BL"){
    col_names <- c("ASV",BL_intersection)
  }else{
    col_names <- c("ASV",EA_intersection)
  }
  dt_022_3_16S <- read.table(paste(PATH,"1_Data/",i,"_022.3_16S_abundances_rareASVandSizeFraction_filtering.tsv", sep=""), header = TRUE, sep = "\t")
  dt_022_3_16S$ASV <- gsub("_16S", "_16S_022_3", dt_022_3_16S$ASV)
  dt_022_3_18S <- read.table(paste(PATH,"1_Data/",i,"_022.3_18S_abundances_rareASVandSizeFraction_filtering.tsv", sep=""), header = TRUE, sep = "\t")
  dt_022_3_18S$ASV <- gsub("_18S", "_18S_022_3", dt_022_3_18S$ASV)
  dt_3_200_16S <- read.table(paste(PATH,"1_Data/",i,"_3.200_16S_abundances_rareASVandSizeFraction_filtering.tsv", sep=""), header = TRUE, sep = "\t")
  dt_3_200_16S$ASV <- gsub("_16S", "_16S_3_200", dt_3_200_16S$ASV)
  dt_3_200_18S <- read.table(paste(PATH,"1_Data/",i,"_3.200_18S_abundances_rareASVandSizeFraction_filtering.tsv", sep=""), header = TRUE, sep = "\t")
  dt_3_200_18S$ASV <- gsub("_18S", "_18S_3_200", dt_3_200_18S$ASV)

  # Write tables
  ## One network per dataset (BL and EA), i.e., we merge kingdom (166S and 18S) and size fraction (022_3 and 3_200)
  write.table(rbind(dt_022_3_16S[,col_names], dt_022_3_18S[,col_names], dt_3_200_16S[,col_names], dt_3_200_18S[,col_names]),
              paste(PATH,"2_Networks/",i,"_ASV_table_for_EnDED.tsv",sep=""), col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
}
```

- prepare ENV data
- reduce ENV table to considered variables
- check if ENV contains all stations
```{r}
# Read in the ENV table
dt <- read.csv(paste(PATH, file_ENV, sep=""), header = TRUE)

# Define the ENV parameters to consider
ENV_IDs <- c("daylength", "Temperature", "Salinity", "NH4", "NO2", "NO3", "PO4", "Si")

# Subset the ENV table to include only relevant columns
if (!all(ENV_IDs %in% colnames(dt))) {
  stop("Error: Some ENV_IDs are not found in the ENV file.")
}

dt <- dt[, c("SampleID", ENV_IDs)]

# Standardize the SampleID column to match the ASV table format
dt$SampleID <- gsub("-", ".", dt$SampleID)

# Transpose the data: rows become columns and columns become rows
dt_t <- t(dt[, -1])  # Exclude SampleID column for transposing
colnames(dt_t) <- dt$SampleID  # Set new column names as SampleID values
dt_t <- as.data.frame(dt_t, stringsAsFactors = FALSE)  # Convert to data frame
dt_t$ENV_ID <- rownames(dt_t)  # Add the ENV_ID column
rownames(dt_t) <- NULL  # Reset row names

# Rearrange columns to have ENV_ID as the first column
dt_t <- dt_t[, c("ENV_ID", setdiff(names(dt_t), "ENV_ID"))]

# Save the transposed table
write.table(
  dt_t, 
  paste(PATH, "1_Data/ENV_allSamples_transposed.tsv", sep=""), 
  col.names = TRUE, 
  row.names = FALSE, 
  sep = "\t", 
  quote = FALSE
)

# BL
# Read in the ASV table
ASV <- read.table(paste(PATH, "2_Networks/BL_ASV_table_for_EnDED.tsv", sep=""), sep="\t", header = TRUE)

# Keep only columns present in the ASV table
common_columns <- intersect(names(dt_t), colnames(ASV))
dt_t_BL <- dt_t[, c("ENV_ID", common_columns)]

# Rename the first column to "ASV"
colnames(dt_t_BL)[1] <- "ASV"
dt_t_BL$ASV <- paste("ENV", dt_t_BL$ASV, sep="_")

# Save table
write.table(
  dt_t_BL, 
  paste(PATH, "2_Networks/BL_ENV.tsv", sep=""), 
  col.names = TRUE, 
  row.names = FALSE, 
  sep = "\t", 
  quote = FALSE
)

# EA
# Read in the ASV table
ASV <- read.table(paste(PATH, "2_Networks/EA_ASV_table_for_EnDED.tsv", sep=""), sep="\t", header = TRUE)

# Keep only columns present in the ASV table
common_columns <- intersect(names(dt_t), colnames(ASV))
dt_t_EA <- dt_t[, c("ENV_ID", common_columns)]

# Rename the first column to "ASV"
colnames(dt_t_EA)[1] <- "ASV"
dt_t_EA$ASV <- paste("ENV", dt_t_EA$ASV, sep="_")

# Save table
write.table(
  dt_t_EA, 
  paste(PATH, "2_Networks/EA_ENV.tsv", sep=""), 
  col.names = TRUE, 
  row.names = FALSE, 
  sep = "\t", 
  quote = FALSE
)
```
-> ENV table contains all needed samples

- prepare ASV-ENV edges
```{r}
ENV <- read.table(paste(PATH,"/1_Data/ENV_allSamples.tsv",sep=""), header = TRUE)

# BL
nodes <- read.table(paste(PATH,"/2_Networks/BL_all_nodes.txt",sep=""), header = FALSE)
ENV_IDs <- c("daylength", "Temperature", "Salinity", "NH4", "NO2", "NO3", "PO4", "Si")
for(e in ENV_IDs)
{
  dt_temp <- data.frame(Source=nodes, Target=paste("ENV", e, sep="_"), association_score=1)
  write.table(dt_temp, paste(PATH,"2_Networks/BL_ASV_",e,".txt",sep=""),
              col.names = TRUE, row.names = FALSE, sep ="\t", quote = FALSE)
}

# EA
nodes <- read.table(paste(PATH,"/2_Networks/EA_all_nodes.txt",sep=""), header = FALSE)
ENV_IDs <- c("daylength", "Temperature", "Salinity", "NH4", "NO2", "NO3", "PO4", "Si")
for(e in ENV_IDs)
{
  dt_temp <- data.frame(Source=nodes, Target=paste("ENV", e, sep="_"), association_score=1)
  write.table(dt_temp, paste(PATH,"2_Networks/EA_ASV_",e,".txt",sep=""),
              col.names = TRUE, row.names = FALSE, sep ="\t", quote = FALSE)
}
```

- EnDED on BL network and EA network
```{bash}
bash EnDED_BL.sh
bash EnDED_EA.sh
```

- BL network: combining EnDED results
```{r}
NW <- read.table(paste(PATH,"/2_Networks/BL_FlashWeave_network.tsv",sep=""), header = TRUE)
NW$original <- 1
dt_temp <- read.table(paste(PATH,"/2_Networks/BL_NW_ENV_daylength.tsv",sep=""), header = TRUE)
NW <- merge(NW, dt_temp[-grep("ENV",dt_temp$Target),c("Source", "Target", "MutualInformation", "InterationInformation", "II_p_value", "DataProcessingInequality_MI_rank", "COMBI_II_DPI")], by=c("Source","Target"), all=TRUE)
colnames(NW)[which(colnames(NW)=="InterationInformation")] <- "ENV_daylength_InterationInformation"
colnames(NW)[which(colnames(NW)=="II_p_value")] <- "ENV_daylength_II_p_value"
colnames(NW)[which(colnames(NW)=="DataProcessingInequality_MI_rank")] <- "ENV_daylength_DataProcessingInequality_MI_rank"
colnames(NW)[which(colnames(NW)=="COMBI_II_DPI")] <- "ENV_daylength_COMBI_II_DPI"

for(i in c("Temperature", "Salinity", "NH4", "NO2", "NO3", "PO4", "Si"))
{
  filename <- paste(PATH,"/2_Networks/BL_NW_ENV_",i,".tsv",sep="")
  dt_temp <- read.table(filename, header = TRUE)
  
  NW <- merge(NW, dt_temp[-grep("ENV",dt_temp$Target),c("Source", "Target", "InterationInformation", "II_p_value", "DataProcessingInequality_MI_rank", "COMBI_II_DPI")], by=c("Source","Target"), all=TRUE)
  colnames(NW)[which(colnames(NW)=="InterationInformation")] <- paste("ENV_",i,"_InterationInformation",sep="")
  colnames(NW)[which(colnames(NW)=="II_p_value")] <- paste("ENV_",i,"_II_p_value",sep="")
  colnames(NW)[which(colnames(NW)=="DataProcessingInequality_MI_rank")] <- paste("ENV_",i,"_DataProcessingInequality_MI_rank",sep="")
  colnames(NW)[which(colnames(NW)=="COMBI_II_DPI")] <- paste("ENV_",i,"_COMBI_II_DPI",sep="")
}

NW$COMBI_allENV <- rowSums(NW[,grep("COMBI",colnames(NW))])
NW_BL <- NW
```

- EA network: combining EnDED results
```{r}
NW <- read.table(paste(PATH,"/2_Networks/EA_FlashWeave_network.tsv",sep=""), header = TRUE)
NW$original <- 1
dt_temp <- read.table(paste(PATH,"/2_Networks/EA_NW_ENV_daylength.tsv",sep=""), header = TRUE)
NW <- merge(NW, dt_temp[-grep("ENV",dt_temp$Target),c("Source", "Target", "MutualInformation", "InterationInformation", "II_p_value", "DataProcessingInequality_MI_rank", "COMBI_II_DPI")], by=c("Source","Target"), all=TRUE)
colnames(NW)[which(colnames(NW)=="InterationInformation")] <- "ENV_daylength_InterationInformation"
colnames(NW)[which(colnames(NW)=="II_p_value")] <- "ENV_daylength_II_p_value"
colnames(NW)[which(colnames(NW)=="DataProcessingInequality_MI_rank")] <- "ENV_daylength_DataProcessingInequality_MI_rank"
colnames(NW)[which(colnames(NW)=="COMBI_II_DPI")] <- "ENV_daylength_COMBI_II_DPI"

for(i in c("Temperature", "Salinity", "NH4", "NO2", "NO3", "PO4", "Si"))
{
  filename <- paste(PATH,"/2_Networks/EA_NW_ENV_",i,".tsv",sep="")
  dt_temp <- read.table(filename, header = TRUE)
  
  NW <- merge(NW, dt_temp[-grep("ENV",dt_temp$Target),c("Source", "Target", "InterationInformation", "II_p_value", "DataProcessingInequality_MI_rank", "COMBI_II_DPI")], by=c("Source","Target"), all=TRUE)
  colnames(NW)[which(colnames(NW)=="InterationInformation")] <- paste("ENV_",i,"_InterationInformation",sep="")
  colnames(NW)[which(colnames(NW)=="II_p_value")] <- paste("ENV_",i,"_II_p_value",sep="")
  colnames(NW)[which(colnames(NW)=="DataProcessingInequality_MI_rank")] <- paste("ENV_",i,"_DataProcessingInequality_MI_rank",sep="")
  colnames(NW)[which(colnames(NW)=="COMBI_II_DPI")] <- paste("ENV_",i,"_COMBI_II_DPI",sep="")
}

NW$COMBI_allENV <- rowSums(NW[,grep("COMBI",colnames(NW))])
NW_EA <- NW
```

- Overview
```{r}
print("BL")
print("all edges")
table(NW_BL$COMBI_allENV)
print("positive edges")
table(NW_BL$COMBI_allENV[which(NW_BL$association_score>0)])
print("negative edges")
table(NW_BL$COMBI_allENV[which(NW_BL$association_score<0)])

print("EA")
print("all edges")
table(NW_EA$COMBI_allENV)
print("positive edges")
table(NW_EA$COMBI_allENV[which(NW_EA$association_score>0)])
print("negative edges")
table(NW_EA$COMBI_allENV[which(NW_EA$association_score<0)])
```
[1] "BL"
[1] "all edges"

   7    8 
   6 2654 
[1] "positive edges"

   7    8 
   2 2393 
[1] "negative edges"

  7   8 
  4 261 

6 edges were kept by 7 ENV, i.e. removed by 1 (2 pos and 4 neg)
2654 edges were kept by 8 ENV, i.e. removed by none (2393 pos and 261 neg)

[1] "EA"
[1] "all edges"

   6    7    8 
   1   13 2421 
[1] "positive edges"

   7    8 
   5 2068 
[1] "negative edges"

  6   7   8 
  1   8 353 

1 edge was kept by 6 ENV, i.e. removed by 2 (0 pos and 1 neg)
13 edges were kept by 7 ENV, i.e. removed by 1 (5 pos and 8 neg)
2421 edges were kept by 8 ENV, i.e. removed by none (2068 pos and 353 neg)

```{r}
print("BL")
paste(length(which(NW_BL$COMBI_allENV==8)), " (", length(which(NW_BL$COMBI_allENV==8))/dim(NW_BL)[1]*100, "%) edges were kept in network")
paste(length(which(NW_BL$COMBI_allENV<8)), " (", length(which(NW_BL$COMBI_allENV<8))/dim(NW_BL)[1]*100, "%) edges were removed from the network")

print("EA")
paste(length(which(NW_EA$COMBI_allENV==8)), " (", length(which(NW_EA$COMBI_allENV==8))/dim(NW_EA)[1]*100, "%) edges were kept in network")
paste(length(which(NW_EA$COMBI_allENV<8)), " (", length(which(NW_EA$COMBI_allENV<8))/dim(NW_EA)[1]*100, "%) edges were removed from the network")
```
[1] "BL"
[1] "2654  ( 99.7744360902256 %) edges were kept in network"
[1] "6  ( 0.225563909774436 %) edges were removed from the network"
[1] "EA"
[1] "2421  ( 99.4250513347023 %) edges were kept in network"
[1] "14  ( 0.574948665297741 %) edges were removed from the network"

Who was responsible for edge removal?
```{r}
print("BL")
for (i in c("daylength", "Temperature", "Salinity", "NH4", "NO2", "NO3", "PO4", "Si")) {
  # Dynamically construct the column name
  column_name <- paste0("ENV_", i, "_COMBI_II_DPI")
  
  # Check if the column exists in NW_BL
  if (column_name %in% colnames(NW_BL)) {
    # Calculate counts and percentages
    zero_count <- length(which(NW_BL[[column_name]] == 0))
    total_rows <- nrow(NW_BL)
    percentage <- (zero_count / total_rows) * 100
    
    # Print the results
    print(paste(
      i, ": ",
      zero_count, 
      " (", round(percentage, 2), "%) ",
      sep = ""
    ))
  } else {
    # Print warning if the column doesn't exist
    print(paste("Column", column_name, "not found in NW_BL"))
  }
}

print("EA")
for (i in c("daylength", "Temperature", "Salinity", "NH4", "NO2", "NO3", "PO4", "Si")) {
  # Dynamically construct the column name
  column_name <- paste0("ENV_", i, "_COMBI_II_DPI")
  
  # Check if the column exists in NW_EA
  if (column_name %in% colnames(NW_EA)) {
    # Calculate counts and percentages
    zero_count <- length(which(NW_EA[[column_name]] == 0))
    total_rows <- nrow(NW_EA)
    percentage <- (zero_count / total_rows) * 100
    
    # Print the results
    print(paste(
      i, ": ",
      zero_count, 
      " (", round(percentage, 2), "%) ",
      sep = ""
    ))
  } else {
    # Print warning if the column doesn't exist
    print(paste("Column", column_name, "not found in NW_EA"))
  }
}
```
[1] "BL"
[1] "daylength: 0 (0%) "
[1] "Temperature: 0 (0%) "
[1] "Salinity: 0 (0%) "
[1] "NH4: 0 (0%) "
[1] "NO2: 3 (0.11%) "
[1] "NO3: 1 (0.04%) "
[1] "PO4: 1 (0.04%) "
[1] "Si: 1 (0.04%) "

[1] "EA"
[1] "daylength: 0 (0%) "
[1] "Temperature: 0 (0%) "
[1] "Salinity: 0 (0%) "
[1] "NH4: 4 (0.16%) "
[1] "NO2: 2 (0.08%) "
[1] "NO3: 2 (0.08%) "
[1] "PO4: 1 (0.04%) "
[1] "Si: 6 (0.25%) "

Filtering out environmentally-driven edges from network
```{r}
# BL
print("BL")
NW <- NW_BL

# Filter the network based on the condition
NW_filtered <- NW[which(NW$COMBI_allENV == 8),]

# Total number of unique nodes
total_nodes <- length(unique(c(as.character(NW$Source), as.character(NW$Target))))
filtered_nodes <- length(unique(c(as.character(NW_filtered$Source), as.character(NW_filtered$Target))))

# Total edges before and after filtering
total_edges_before <- dim(NW)[1]
total_edges_after <- dim(NW_filtered)[1]

# Positive and negative edge counts before filtering
positive_before <- length(which(NW$association_score > 0))
negative_before <- length(which(NW$association_score < 0))
positive_before_percent <- (positive_before / total_edges_before) * 100
negative_before_percent <- (negative_before / total_edges_before) * 100

# Positive and negative edge counts after filtering
positive_after <- length(which(NW_filtered$association_score > 0))
negative_after <- length(which(NW_filtered$association_score < 0))
positive_after_percent <- (positive_after / total_edges_after) * 100
negative_after_percent <- (negative_after / total_edges_after) * 100

# Print the results in the required format
cat("-> Before EnDED:", total_nodes, "nodes\n")
cat("-> After EnDED:", filtered_nodes, "nodes\n")
cat("-> Before EnDED: ", total_edges_before, " edges (", positive_before, ", ", 
    round(positive_before_percent, 2), "% positive, and ", 
    negative_before, ", ", round(negative_before_percent, 2), "% negative)\n", sep = "")
cat("-> After EnDED: ", total_edges_after, " edges (", positive_after, ", ", 
    round(positive_after_percent, 2), "% positive, and ", 
    negative_after, ", ", round(negative_after_percent, 2), "% negative)\n", sep = "")

NW_filtered_BL <- NW_filtered

# EA
print("EA")
NW <- NW_EA

# Filter the network based on the condition
NW_filtered <- NW[which(NW$COMBI_allENV == 8),]

# Total number of unique nodes
total_nodes <- length(unique(c(as.character(NW$Source), as.character(NW$Target))))
filtered_nodes <- length(unique(c(as.character(NW_filtered$Source), as.character(NW_filtered$Target))))

# Total edges before and after filtering
total_edges_before <- dim(NW)[1]
total_edges_after <- dim(NW_filtered)[1]

# Positive and negative edge counts before filtering
positive_before <- length(which(NW$association_score > 0))
negative_before <- length(which(NW$association_score < 0))
positive_before_percent <- (positive_before / total_edges_before) * 100
negative_before_percent <- (negative_before / total_edges_before) * 100

# Positive and negative edge counts after filtering
positive_after <- length(which(NW_filtered$association_score > 0))
negative_after <- length(which(NW_filtered$association_score < 0))
positive_after_percent <- (positive_after / total_edges_after) * 100
negative_after_percent <- (negative_after / total_edges_after) * 100

# Print the results in the required format
cat("-> Before EnDED:", total_nodes, "nodes\n")
cat("-> After EnDED:", filtered_nodes, "nodes\n")
cat("-> Before EnDED: ", total_edges_before, " edges (", positive_before, ", ", 
    round(positive_before_percent, 2), "% positive, and ", 
    negative_before, ", ", round(negative_before_percent, 2), "% negative)\n", sep = "")
cat("-> After EnDED: ", total_edges_after, " edges (", positive_after, ", ", 
    round(positive_after_percent, 2), "% positive, and ", 
    negative_after, ", ", round(negative_after_percent, 2), "% negative)\n", sep = "")

NW_filtered_EA <- NW_filtered
```
[1] "BL"
-> Before EnDED: 1914 nodes
-> After EnDED: 1913 nodes
-> Before EnDED: 2660 edges (2395, 90.04% positive, and 265, 9.96% negative)
-> After EnDED: 2654 edges (2393, 90.17% positive, and 261, 9.83% negative)

[1] "EA"
-> Before EnDED: 2099 nodes
-> After EnDED: 2091 nodes
-> Before EnDED: 2435 edges (2073, 85.13% positive, and 362, 14.87% negative)
-> After EnDED: 2421 edges (2068, 85.42% positive, and 353, 14.58% negative)

Save tables
```{r}
write.table(NW_BL,
            paste(PATH,"/2_Networks/BL_NW_withEnDED.tsv",sep=""),
            col.names = TRUE, row.names = FALSE, sep = "\t", dec = ".", quote = FALSE)
write.table(NW_filtered_BL[,c("Source","Target","association_score","MutualInformation")],
            paste(PATH,"/2_Networks/BL_NW_afterEnDED.tsv",sep=""),
            col.names = TRUE, row.names = FALSE, sep = "\t", dec = ".", quote = FALSE)

write.table(NW_EA,
            paste(PATH,"/2_Networks/EA_NW_withEnDED.tsv",sep=""),
            col.names = TRUE, row.names = FALSE, sep = "\t", dec = ".", quote = FALSE)
write.table(NW_filtered_EA[,c("Source","Target","association_score","MutualInformation")],
            paste(PATH,"/2_Networks/EA_NW_afterEnDED.tsv",sep=""),
            col.names = TRUE, row.names = FALSE, sep = "\t", dec = ".", quote = FALSE)
```

## Co-occurrence filter
```{r}
# BL
NW_BL <- read.table(paste(PATH,"/2_Networks/BL_NW_afterEnDED.tsv",sep=""), header = TRUE, sep = "\t")
coOcc <- read.table(paste(PATH,"/2_Networks/BL_NW_CoOcc.tsv",sep=""), header = TRUE, sep = "\t")
NW_BL <- merge(NW_BL, coOcc[,-3], by=c("Source","Target"), all.x=TRUE, all.y=FALSE)

# EA
NW_EA <- read.table(paste(PATH,"/2_Networks/EA_NW_afterEnDED.tsv",sep=""), header = TRUE, sep = "\t")
coOcc <- read.table(paste(PATH,"/2_Networks/EA_NW_CoOcc.tsv",sep=""), header = TRUE, sep = "\t")
NW_EA <- merge(NW_EA, coOcc[,-3], by=c("Source","Target"), all.x=TRUE, all.y=FALSE)
```

```{r}
print("BL")
NW <- NW_BL
cat(
  "Nodes with percentage_co_occurrence > 0.5:", 
  length(unique(c(as.character(NW[which(NW$percentage_co_occurrence > 0.5), "Source"]), as.character(NW[which(NW$percentage_co_occurrence > 0.5), "Target"])))), 
  "| All edges:", 
  length(which(NW$percentage_co_occurrence <= 0.5)), "with percentage_co_occurrence <= 0.5,", 
  length(which(NW$percentage_co_occurrence > 0.5)), "with percentage_co_occurrence > 0.5 | Positive edges:", 
  length(which(NW$percentage_co_occurrence <= 0.5 & NW$association_score > 0)), "with percentage_co_occurrence <= 0.5,", 
  length(which(NW$percentage_co_occurrence > 0.5 & NW$association_score > 0)), "with percentage_co_occurrence > 0.5 | Negative edges:", 
  length(which(NW$percentage_co_occurrence <= 0.5 & NW$association_score < 0)), "with percentage_co_occurrence <= 0.5,", 
  length(which(NW$percentage_co_occurrence > 0.5 & NW$association_score < 0)), "with percentage_co_occurrence > 0.5 | % Positive with percentage_co_occurrence > 0.5:", 
  round(length(which(NW$percentage_co_occurrence > 0.5 & NW$association_score > 0)) / length(which(NW$percentage_co_occurrence > 0.5)) * 100, 2), 
  "% | % Negative with percentage_co_occurrence > 0.5:", 
  round(length(which(NW$percentage_co_occurrence > 0.5 & NW$association_score < 0)) / length(which(NW$percentage_co_occurrence > 0.5)) * 100, 2), 
  "%\n"
)

print("EA")
NW <- NW_EA
cat(
  "Nodes with percentage_co_occurrence > 0.5:", 
  length(unique(c(as.character(NW[which(NW$percentage_co_occurrence > 0.5), "Source"]), as.character(NW[which(NW$percentage_co_occurrence > 0.5), "Target"])))), 
  "| All edges:", 
  length(which(NW$percentage_co_occurrence <= 0.5)), "with percentage_co_occurrence <= 0.5,", 
  length(which(NW$percentage_co_occurrence > 0.5)), "with percentage_co_occurrence > 0.5 | Positive edges:", 
  length(which(NW$percentage_co_occurrence <= 0.5 & NW$association_score > 0)), "with percentage_co_occurrence <= 0.5,", 
  length(which(NW$percentage_co_occurrence > 0.5 & NW$association_score > 0)), "with percentage_co_occurrence > 0.5 | Negative edges:", 
  length(which(NW$percentage_co_occurrence <= 0.5 & NW$association_score < 0)), "with percentage_co_occurrence <= 0.5,", 
  length(which(NW$percentage_co_occurrence > 0.5 & NW$association_score < 0)), "with percentage_co_occurrence > 0.5 | % Positive with percentage_co_occurrence > 0.5:", 
  round(length(which(NW$percentage_co_occurrence > 0.5 & NW$association_score > 0)) / length(which(NW$percentage_co_occurrence > 0.5)) * 100, 2), 
  "% | % Negative with percentage_co_occurrence > 0.5:", 
  round(length(which(NW$percentage_co_occurrence > 0.5 & NW$association_score < 0)) / length(which(NW$percentage_co_occurrence > 0.5)) * 100, 2), 
  "%\n"
)
```
[1] "BL"
Nodes with percentage_co_occurrence > 0.5: 1906
All edges: 40 with percentage_co_occurrence <= 0.5, 2614 with percentage_co_occurrence > 0.5
Positive edges: 0 with percentage_co_occurrence <= 0.5, 2393 with percentage_co_occurrence > 0.5
Negative edges: 40 with percentage_co_occurrence <= 0.5, 221 with percentage_co_occurrence > 0.5
% Positive with percentage_co_occurrence > 0.5: 91.55 %
% Negative with percentage_co_occurrence > 0.5: 8.45 %

[1] "EA"
Nodes with percentage_co_occurrence > 0.5: 2074
All edges: 32 with percentage_co_occurrence <= 0.5, 2389 with percentage_co_occurrence > 0.5
Positive edges: 0 with percentage_co_occurrence <= 0.5, 2068 with percentage_co_occurrence > 0.5
Negative edges: 32 with percentage_co_occurrence <= 0.5, 321 with percentage_co_occurrence > 0.5
% Positive with percentage_co_occurrence > 0.5: 86.56 %
% Negative with percentage_co_occurrence > 0.5: 13.44 %

```{r}
# BL
write.table(NW_BL[which(NW_BL$percentage_co_occurrence>0.5),],
            paste(PATH,"/2_Networks/BL_NW_static.tsv",sep=""),
            col.names = TRUE, row.names = FALSE, sep = "\t", dec = ".", quote = FALSE)

# EA
write.table(NW_EA[which(NW_EA$percentage_co_occurrence>0.5),],
            paste(PATH,"/2_Networks/EA_NW_static.tsv",sep=""),
            col.names = TRUE, row.names = FALSE, sep = "\t", dec = ".", quote = FALSE)
```

### 2.3. Determining temporal network
- determine temporal network from single static network via subnetworks
- an edge is present in a monthly subnetwork when both microorganisms of the predicted associations are present (abundance > 0)
```{r}
#########################################################
# BL
#########################################################
dt <- read.table(paste(PATH,"2_Networks/BL_ASV_table_for_EnDED.tsv", sep=""), header = TRUE, sep = "\t")
dt[is.na(dt)] <- 0 # Replace all NA values in the data frame with 0
global_min <- 0
rownames(dt) <- dt$ASV
dt$ASV <- NULL
NW <- read.table(paste(PATH,"2_Networks/BL_NW_static.tsv", sep=""), header = TRUE, sep = "\t")
colnames(NW)[1:3] <- c("Source", "Target", "Weight")
#hist(NW$Weight, breaks = 100)
for(i in colnames(dt)){
  NW[,i] <- 1*((dt[NW$Source,i]>global_min) & (dt[NW$Target,i]>global_min))
}
write.table(NW, paste(PATH,"2_Networks/BL_temporalNetwork.tsv", sep=""), col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)

#########################################################
# EA
#########################################################
dt <- read.table(paste(PATH,"2_Networks/EA_ASV_table_for_EnDED.tsv", sep=""), header = TRUE, sep = "\t")
dt[is.na(dt)] <- 0 # Replace all NA values in the data frame with 0
global_min <- 0
rownames(dt) <- dt$ASV
dt$ASV <- NULL
NW <- read.table(paste(PATH,"2_Networks/EA_NW_static.tsv", sep=""), header = TRUE, sep = "\t")
colnames(NW)[1:3] <- c("Source", "Target", "Weight")
#hist(NW$Weight, breaks = 100)
for(i in colnames(dt)){
  NW[,i] <- 1*((dt[NW$Source,i]>global_min) & (dt[NW$Target,i]>global_min))
}
write.table(NW, paste(PATH,"2_Networks/EA_temporalNetwork.tsv", sep=""), col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE)
```
